\documentclass[10pt,a4paper]{article}
\usepackage{url}

\pagestyle{empty}

\author{Armijn Hemel -- Binary Analysis Tool project}
\title{Binary Analysis Tool User and Developer Manual - describing version 36}

\begin{document}

\maketitle
\thispagestyle{empty}

\tableofcontents

\section{Introducing the Binary Analysis Tool}

The Binary Analysis Tool (BAT) is a generic framework that can help developers
and companies analyse binary files. Its primary application is for Open Source
software license compliance, with a special focus on supply chain management
in consumer electronics, but it can also be used for other checks of binary
for example the presence of security bugs.

BAT consists of several programs written in Python. The most important program
is the scanner for binary objects to unpack binaries recursively and apply a
number of scans, for example for open source license compliance, visualising
linking information, finding version information, and so on. There are also
other programs to help with specific license compliance tasks, such as
verifying if configurations for a given BusyBox binary match with the
configuration in source code. Also included is a very experimental program to
derive a possible configuration from a Linux kernel image, as well as programs
to verify results from a binary scan with a source code archive.

Development of security analysis features in BAT has been made possible through
a joint grant from NLnet foundation and the programme ``veilig door innovatie''
from NCTV.

\section{Installing the Binary Analysis Tool}

\subsection{Hardware requirements}

The tools in the Binary Analysis Tool can be quite resource intensive. They
are largely I/O-bound (database access, reading files from disk), so it is
better to invest in faster disks or ramdisks than in raw CPU power. Using more
cores is also highly recommended, since most of the programs in the Binary
Analysis Tool can take advantage of this and will run significantly faster.

\subsection{Software requirements}

To run BAT a recent Linux distribution is needed, such as a recent Fedora or
Ubuntu. Ubuntu versions older than 14.04, will not work due to a broken
version of the PyDot package (unless a newer version is installed using
\texttt{pip}). Debian versions older than 7 are unsupported.

Versions older than Fedora 20 might not work scanning because of a bug in the
version of matplotlib shipped on those distributions will throw errors in some
cases.

If the latest version from version control is used it is important to look at
the file \texttt{setup.cfg} to get a list of the dependencies that should be
met on the host system before installing BAT if the host system is Fedora. If
the host system is Ubuntu or Debian this information will be in
\texttt{debian/control}.

\subsubsection{Security warning}

Do not install BAT on a machine that is performing any critical functions for
your organisation. There are certain pieces of code in BAT that have known
security issues, such as some of the Squashfs unpacking programs in
\texttt{bat-extratools} that have been lifted from vendor SDKs. When scanning
untrusted binary code there might be security risk.

\subsubsection{Installation on Fedora}

To install on Fedora two packages are needed: \texttt{bat-extratools}, %, \texttt{ubi\_reader}
and \texttt{bat}. These packages easily be created as both binary versions and
as source RPM files from the source code in Git (as described later). When
installing the two files there should be a list of dependencies that should be
installed to let BAT work successfully. Some of the dependencies are not in
Fedora by default but need to be installed through external repositories, such
as RPMfusion.

\subsubsection{Installation on Debian and Ubuntu}

To install on Debian and Ubuntu two packages are needed:
\texttt{bat-extratools}, %, \texttt{ubi\_reader}
and \texttt{bat}. These can easily be generated as DEB packages using the
commands described later in this manual.

When installing the two files there should be a list of dependencies
that should be installed to let BAT work successfully. Some of these packages
are not in Debian by default but need to be installed by enabling extra
repositories such as Debian \texttt{non-free}.

\subsubsection{Installation on CentOS}

In some cases it is possible to run BAT on CentOS (6.7 or 7 has been tested
with) but some functionality will not be available, such as UBI/UBIFS
unpacking and the scans creating graphs with PyDot (ELF linking, kernel module
linking).

It might be necessary to enable the EPEL repository (
\url{https://fedoraproject.org/wiki/EPEL} ) as well as RepoForge. A few
packages might have to be installed manually.

\section{Analysing binaries with the Binary Analysis Tool}

BAT consists of several programs and a few helper scripts (not meant to be
used directly). The main purpose of the Binary Analysis tool is to analyse
arbitrary binaries and review results. Analysis of the binary is done via
a commandline tool (\texttt{bat-scan}), while the results can be viewed using
a special graphical viewer (\texttt{batgui}).

\subsection{Running \texttt{bat-scan}}

The \texttt{bat-scan} program can scan in two modes: either scan a single
binary, or scan a whole directory of files. These are mutually exclusive and
you cannot mix and match parameters for both modes.

To scan a single binary you will need to supply three parameters to
\texttt{bat-scan}:

\begin{enumerate}
\item \texttt{-c} : path to a configuration file
\item \texttt{-b} : path to the binary to be scanned
\item \texttt{-o} : path to an output file, where unpacked files, reports,
plus the final program state be written to. This file can later be opened with
the viewer.
\end{enumerate}

The default install of BAT comes with a configuration file (installed in
\texttt{/etc/bat/} although this will likely change in the future) with default
settings that have proven to work well but almost everything can be changed or
tweaked. A lengthy explanation of the different types of scans and their
configuration can be found in the appendix.

A typical invocation looks like this:

\begin{verbatim}
python bat-scan -c /path/to/configuration -b /path/to/binary -o
/path/to/outputfile
\end{verbatim}

To scan a directory you will need to supply three parameters to
\texttt{bat-scan}:

\begin{enumerate}
\item \texttt{-c} : path to a configuration file
\item \texttt{-d} : path to a directory with files to be scanned
\item \texttt{-u} : path to a directory where output files will be written to
\end{enumerate}

For example:

\begin{verbatim}
python bat-scan -c /path/to/configuration -d /path/to/dirwithbinaries
-u /path/to/dirwithoutputfiles
\end{verbatim}

The format of output files in ``directory scan'' mode will be the name of the
original file with the suffix \texttt{.tar.gz}. If there is already a file with
that name in the output directory the file will not be scanned again. If the
file should be scanned again, then the output file should be (re)moved.

\subsection{Interpreting the results}

\texttt{bat-scan} will output an rchive file containing program state, complete
unpacked directory tree containing all unpacked data (unless \texttt{outputlite}
was set to \texttt{yes}), plus possibly some extra generated data, such as
pictures and more reporting. These dumps are meant to be used by
\texttt{batgui} or processed by other programs.

\subsubsection{Output archive}

The output archive contains a few files and directories (depending on scan
configuration):

\begin{itemize}
\item \texttt{scandata.pickle} - Python pickle containing information about the
structure of the binary, including offsets, paths, tags, and so on. It does not
contain any of the actual scan results.
\item \texttt{scandata.json} - JSON file containing a subset of
the information in \texttt{scandata.pickle}. This file is only generated if the
\texttt{generatejson} scan is enabled.
\item \texttt{STATISTICS} - text file containing some statistics about the
version of BAT used, the underlying Python implementation, and the scan times
for each of the phases and some subphases. This file will likely be replaced
by a JSON file in the future.
\item \texttt{data} - directory containing the full unpacked directory tree.
If \texttt{outputlite} is set to \texttt{yes} this directory will be omitted
from the output archive.
\item \texttt{filereports} - directory containing Python pickle files (gzip
compressed) with scan results. Since identical files might be present the
results are stored per checksum, not file name.
\item \texttt{images} - directory containing various images with results of
scans (depending on which scans are enabled), per checksum
\item \texttt{offsets} - directory with Python pickle files containing the
offsets of possible file systems, compressed files and media files found in
the file. This directory as well as its files will only be created if
\texttt{dumpoffsets} is set to \texttt{yes} in the global configuration.
\item \texttt{reports} - directory containing HTML and (optionally) JSON
reports, per checksum
\end{itemize}

\subsubsection{Viewing results with \texttt{batgui}}

The \texttt{batgui} program was made to view the results of the analysis
process easily. The viewer has two modes: simple and advanced. In simple mode
a tree of the unpack results will be shown, and each file in the tree can be
clicked to display more information. Depending on which scans were run the tree
will be decorated with more information, such as the type of the file (based on
tags), or if matches were found with the ranking method. Using a filtering
system (available from the menu) files that are typically uninteresting for
license compliance engineering (empty files/directories, symbolic links,
graphics files and so on) can be ignored.

Information that is shown per file depends on the scans that were run and the
type of file. For most files information like size, type, path (both relative
inside the unpacked binary, as well as absolute in the scanning tree) will be
shown. If the ranking method was enabled results of the ranking process such as
matched strings, function names, a license guess etecetera will be displayed as
well.

In the optional advanced mode more results will be shown, such as a graphical
representation of a file, where every bit in the binary has been assigned a
grayscale value, plus a textual representation of a file generated with
\texttt{hexdump}. Advanced mode is disabled by default, since loading the
additional pictures and data is quite resource intensive and it will only be
useful in very specific cases. It also requires that these special files are
generated by BAT when scanning a file. This is not done by default but needs
an explicit configuration change. Advanced mode might be removed from the GUI
in future versions of BAT.

The \texttt{batgui} program is no longer part of the default distribution of
BAT, but can be downloaded from the BAT repository.

\subsubsection{Viewing results with \texttt{batgui2}}

A new user interface was created, based on Qt5. It is not part of the regular
distribution of BAT, but can be grabbed from
\url{https://github.com/monkeyiq/batgui2}

\section{Additional programs in the Binary Analysis Tool}

\subsection{\texttt{busybox.py} and \texttt{busybox-compare-configs.py}}

Two other tools in BAT are \texttt{busybox-compare-configs.py} and
\texttt{busybox.py} (in the subdirectory \texttt{bat}). These two tools are
specifically used to analyse BusyBox binaries. BusyBox is in widespread use on
embedded devices and the license violations of BusyBox are actively enforced
in court.

BusyBox binaries on embedded machines often have different configurations,
depending on the needs of the manufacturer. Since providing the correct
configuration is one of the requirements for license compliance it is important
to be able to determine the configuration of a BusyBox binary and verify that
there is a corresponding configuration file in the source code release.

The BusyBox processing tools in BAT try to extract the most likely
configuration from the binary and print it in the right format for that version
of BusyBox.

\texttt{busybox.py} is used to extract the configuration from a binary.
Afterwards \texttt{busybox-compare-configs.py} can be used to compare the
extracted configuration with a vendor supplied configuration.

\subsubsection{Extracting a configuration from BusyBox}

Extracting a configuration from a BusyBox executable is done using
\texttt{busybox.py} which can be found in the \texttt{bat} directory. It needs
two commandline parameters: the path to the binary and the path to a directory
which has files containing mappings from BusyBox applet names to BusyBox
configuration directives. By default this value is hardcoded as
\texttt{/etc/bat}, but this might change in the future. Some pre-extracted
configurations can be found in the \texttt{bat-data} package (coming soon).

The output (a possible BusyBox configuration) is written to standard output.

\begin{verbatim}
python bat/busybox.py -b /path/to/busybox/binary -c
     /path/to/pre/extracted/configs > /path/to/saved/config
\end{verbatim}

This command will save the configuration to a file, which can be used as an
input to \texttt{busybox-compare-configs.py}.

\subsubsection{Comparing two BusyBox configurations}

After extracting the configuration the extracted configuration can be compared
to another configuration, for example a configuration as supplied by a vendor
in a source code archive:

\begin{verbatim}
python busybox-compare-configs.py -e /path/to/saved/config
    -f /path/to/vendor/configuration -n $version
\end{verbatim}

\subsection{\texttt{comparebinaries.py}}

The \texttt{comparebinaries.py} program compares two file trees with for example
unpacked firmwares. It is intended to find out which differences there are
between two binaries (like firmwares) unpacked with BAT.

There are two scenarios where this program can be used:

\begin{enumerate}
\item comparing an old firmware (that is already known and which has been
verified) to a new firmware (update) and see if there are any differences.
\item comparing a firmware to a rebuild of a firmware as part of compliance
engineering.
\end{enumerate}

A few assumptions are made:

\begin{enumerate}
\item both firmwares were unpacked using the Binary Analysis Tool
\item files that are in the original firmware, but not in the new firmware, are
not reported (example: removed binaries). This will change in a future version.
\item files that are in the new firmware but not not in the original firmware are
reported, since this would mean additions to the firmware which need to be
checked.
\item files that appear in both firmwares but which are not identical are
checked using \texttt{bsdiff} to determine the size of the difference.
\end{enumerate}

With checksums it is easy to find the files that are different. Using
\texttt{bsdiff} it becomes easier to prioritise based on the size of the
difference.

Small differences are probably not very interesting at all:

\begin{enumerate}
\item time stamps (BusyBox, Linux kernel, and others record a time stamp in the
binary)
\item slightly different build system settings (home directories, paths, and
so on).
\end{enumerate}

Bigger differences are of course much more interesting.

\subsection{\texttt{sourcewalk.py}}

This program can quickly determine whether or not source code files in a
directory can be found in known upstream sources. It uses a pregenerated
database containing names and checksums of files (for example the Linux kernel)
and reports whether or not the source code files can be found in the database
based on these checksums.

The purpose of this script is to find source code files that cannot be found in
upstream sources to reduce the search space during a source code audit.

This script will not catch:

\begin{itemize}
\item binary files
\item patch/diff files
\item anything that does not have an extension from the list in the script
\item configuration files/build scripts
\end{itemize}

\subsection{\texttt{verifysourcearchive.py}}

The \texttt{verifysourcearchive.py} program is to verify a source code archive
using the result of a scan done with BAT.

\subsection{\texttt{findxor.py}}

The \texttt{findxor.py} program can be used to find possible XOR ``encryption''
keys. It prints the top 10 (hardcoded limit) of most common byte sequences (16
bytes) in the file. These can then be added to the \texttt{batxor.py} module in
BAT. This will likely change in the future.

\section{Binary Analysis Tool extratools collection}

To help with unpacking non-standard file systems, or standard file systems for
which there are no tools readily available on Fedora or Ubuntu there is also
a collection of tools that can be used by BAT to unpack more file systems.
These tools are not part of the standard distribution, but have to be installed
separately. They are governed by different license conditions than the core BAT
distribution.

Currently the collection consists of:

\begin{itemize}
\item \texttt{bat-minix} has a Python script to unpack Minix v1 file systems
that are frequently found on older embedded Linux systems, such as IP cameras.
\item modified version of \texttt{code2html} (which is unmaintained by the
upstream author) that adds support for various more programming languages.
This tool is not needed by BAT.
\item unmodified version of \texttt{simg2img} needed for converting Android
sparse files to ext4 file system images.
\item unmodified version of \texttt{romfsck} needed for unpacking romfs file
systems.
\item modified version of \texttt{cramfsck} that enables unpacking cramfs
file systems.
\item reimplementation of \texttt{unyaffs} that enables unpacking for various
YAFFS2 file systems.
\item various versions of \texttt{unsquashfs} that enable unpacking variants
of SquashFS. These versions have either been lifted from vendor SDKs, the
OpenWrt project, DD-WRT, or upstream SquashFS project.
\item \texttt{ubi\_reader} is a set of tools to deal with UBI/UBIFS images,
currently not used by default.
\item \texttt{bat-visualisation} containing a few custom tools to help generate
pictures. These might be removed in the future.
\end{itemize}

The collection is split in two packages: the \texttt{ubi\_reader} package contains UBI/UBIFS
specific tools and the \texttt{bat-extratools} package contains the rest.

\appendix

\section{BAT scanning phases}

BAT uses a brute force approach for analysing a binary. It assumes no prior
knowledge of how a binary is constructed or what is inside the binary. Instead
it tries to determine what is inside by applying a wide range of methods, such
as looking for known identifiers of file systems and compressed files and
running external tools to find contents in the binary. It should be noted that
there are possibilities to add more information to the system to speed up
scanning and skip phases.

During scanning of a file the following steps are taken:

\begin{enumerate}
\item identifier search, using a list of known identifiers, like headers,
footers or identifiers that indicate the start or end of a file system,
compressed file or media file.
\item verifying file type of a file and, if successful, tagging it. Tags can
be used later on to give more information to the scanner.
\item unpacking file systems, compressed files and media files from the file,
carving them out of the larger file first.
\item repeat steps 1 - 3 for each file that was unpacked in step 3
\item run individual scans on each file if no further unpacking is possible
\item optionally aggregate scan results or modify results based on information
that has become available during the scan
\item process results from scans in step 5 and 6 and generate reports
\item pack results into an archive that can be used by the viewer application
or other applications
\end{enumerate}

\subsection{Identifier search}

The first action performed is scanning a file for known identifiers of
compressed files, file systems and media files. The identifers are important
for a few reasons: first, they are used to determine which checks will run. They
are also used frequently throughout the code for verification and speeding up
unpacking. If a scan depends on a specific identifier being present it can be
set using the \texttt{magic} attribute in the configuration. If an identifier
is not defined anywhere in the configuration file as needed it will be skipped
during the identifier search to speed up the identifier search. Some scans
define an additional magic header in \texttt{optmagic}. The values defined in
\texttt{optmagic} are not authoritive, but should be treated as hints. A good
example is the YAFFS2 scan.

The marker search cannot be enabled or disabled via the configuration file. The
markers that are searched for can be found in \url{bat/fsmagic.py}.

As an optimization the marker search can be skipped for some files if they
have an extension which gives a possible hint about what kind of file it might
be. For example, the extension \texttt{gz} is frequently used for gzip
compressed files, so for files with the extension \texttt{gz} a special method
(configured in the configuration for the gzip unpacker) is first run to see if
the whole file is actually a gzip file, without looking at any other markers,
or trying other scans first. If the whole file is indeed gzip compressed (which
will be the case for the vast majority of files) then all other unpacking scans
will be skipped. If the file is not gzip compressed, or only part of the file
is gzip compressed (and there is trailing data), then the file will be
processed in the normal way instead.

If multiple CPUs are available and the top level file is larger than a certain
limit and does not have a known extension as described above the marker search
will be done in parallel as a speed up. The limit can be set in the global
configuration using the variable \texttt{markersearchminimum}. The default
value for this variable is 20 million bytes.

\subsection{Pre-run checks}

Before files are unpacked they are briefly inspected and if possible tagged.
Tags are used to pass hints to methods that are run later to avoid unnecessarily
scanning a file and to reduce the amount of false positives.

For example, files that only contain text are tagged as \texttt{text}, all other
files are tagged as \texttt{binary} (this depends on the implementation of
Python. Python 2 only considers (by default) ASCII to be valid text). Methods
that only work on binaries can then ignore anything that has been tagged as
\texttt{text}.

Other checks that are available are for valid XML, various Android formats,
ELF executables and libraries, certain graphics and audio files, and so on.

The prerun checks can easily be identified in the configuration, since it has
its type set to \texttt{prerun}:

\begin{verbatim}
[verifytext]
type        = prerun
module      = bat.prerun
method      = verifyText
priority    = 3
description = Check if file contains just ASCII text
enabled     = yes
\end{verbatim}

Prerun verifiers can optionally make use of tags that are already present by
using \texttt{magic} and \texttt{noscan} attributes, which will be explained
in detail later for the unpackers.

\subsection{Unpackers}

Unpackers can be recognized in the configuration because their type is set
to \texttt{unpack}, for example:

\begin{verbatim}
[jffs2]
type        = unpack
module      = bat.fwunpack
method      = searchUnpackJffs2
priority    = 2
magic       = jffs2_le:jffs2_be
noscan      = text:xml:graphics:pdf:compressed:audio:video:mp4:elf:java:resource:dalvik
description = Unpack JFFS2 file systems
enabled     = yes
\end{verbatim}

In BAT 36 the following file systems, compressed files and media files can be
unpacked or extracted:

\begin{itemize}
\item file systems: Android sparse files, cramfs, ext2/ext3/ext4, ISO9660,
JFFS2 (no LZO compression), Minix (specific variant of v1 often found on older
embedded Linux systems), SquashFS (several variants), romfs, YAFFS2 (specific
variants), ubifs (not on all systems), PLF (Parrot specific file format)
\item compressed files and executable formats: 7z, ar, ARJ, BASE64, BZIP2,
compressed Flash, CAB, compress, CPIO, ELF, EXE (specific compression methods only),
GZIP, InstallShield (old versions), Java class file, LRZIP, LZIP, LZMA, LZOP, MSI, pack200, RAR,
RPM, RZIP, serialized Java, TAR, UPX, XZ, ZIP (including APK, EAR, JAR and
WAR), WIM, Intel HEX (whole files only, comments allowed), XAR
\item media files: BMP, GIF, JPEG, PNG, ICO, PDF, CHM, OTF, TTF, WOFF
\end{itemize}

Most of the unpackers for these file systems, compressed files and media files
are located in the file \url{bat/fwunpack.py}.

Unpacking differs per file type. Most files use one or more identifiers that
can be searched for in a binary blob. Using this information it is possible
to carve out the right parts of a binary blob and verify if it indeed contains
a compressed file, media file or file system.

There is not always an identifier that can be searched for. The YAFFS2 file
system layout for example is dependent on the hardware specifics of the
underlying flash chip. Without knowing these specifics it is not possible to
specifically search for a valid YAFFS2 file system. This scan therefore tries
to run on every file, unless explicitely filtered out (using \texttt{noscan}
and tags).

Other file types (such as ARJ files) have a very generic identifier, so there
are a lot of false positives. This causes a big increase in runtime. The ARJ
unpacker is therefore disabled by default.

LZMA is another special case: there are many different valid headers for LZMA
files, but in practice only a handful are used.

If unpacking is successful a directory with unpacked files is returned, and, if
available, some meta information to avoid duplicate scanning (blacklisting
information and tags). The unpacked files are added to the scan queue and
scanned recursively.

\subsection{Leaf scans}

Leaf scans are scans that are run on every single file after unpacking,
including files that contained files that were found and extracted by unpackers.

Leaf scans can be recognized in the configuration because their type is set to
\texttt{leaf}, for example:

\begin{verbatim}
[markers]
type        = leaf
module      = bat.checks
method      = searchMarker
noscan      = text:xml:graphics:pdf:compressed:audio:video
description = Determine presence of markers of several open source programs
enabled     = yes
\end{verbatim}

The current leaf scans that are available in BAT are:

\begin{itemize}
\item marker scan searching for signature scans of a few open source programs
(dproxy, ez-ipupdate, hostapd, iptables, iproute, libusb, loadlin, RedBoot,
U-Boot, vsftpd, wireless-tools, wpa-supplicant)
\item advanced search mode using ranking of strings, function names, variable
names, field names and Java class names using a database (for ELF and Java,
both regular JVM and Dalvik)
\item BusyBox version number
\item dynamic library dependencies (ELF files only)
\item file architecture (ELF files only)
\item Linux kernel module license (Linux kernel modules only)
\item Linux kernel version number, plus detection for several subsystems
\item PDF meta data extraction
\item presence of URLs indicating an open source license
\item presence of URLs indicating forges/collaborative software development
sites (SourceForge, GitHub, etcetera)
\end{itemize}

The fast string searches are meant for quick sweep scanning only. They have
their limits, can report false positives or fail to identify a binary. They
should only be used to signal that further inspection is necessary. For a
thorough investigation the advanced search mode should be used. These scans are
likely to be disabled in the future in the default configuration.

\subsection{Aggregators}

Sometimes it helps to aggregate results of a number of files, or it could be
useful to perform other actions after all the individual scans have run. The
best example is dealing with JAR-files (Java ARchives). Individual Java class
files often contain too little information to map them reliably to a source
code package.

Typically a class file contains just a few method names, or field names, or
strings. If inner classes are used it can be even worse and information from a
single source code file could be scattered across several class files.

Since Java programs (note: excluding Android) are typically distributed as a
JAR that is either included at runtime or directly executed, similar to an ELF
library or ELF executable, it makes perfect sense to treat the JAR file as a
single unit and aggregate results for the individual class files and assign
them to the JAR file.

Aggregators take all results of the entire scan as input.

Currently the following aggregators are available:

\begin{itemize}
\item advanced identifier search and classification
\item aggregating result of individual Java class files in case they come from
the same JAR file.
\item cleaning up/fixing results of duplicate files: often firmwares contain
duplicate files. Sometimes some more information is available to make a better
choice as to which file is the duplicate and which one is the original version
\item checking dynamically linked ELF files
\item finding duplicate files
\item finding licenses and versions of strings and function names that were
found and optionally pruning the result set to remove unlikely results.
\item pruning files from the scan completely if they are not interesting (such
as pictures, or text files) using tags.
\item generating pictures of results of a scan
\item generating reports of results of a scan
\end{itemize}

\subsection{Post-run methods}

In BAT there are methods that are run after all the regular work has been
performed, or ``post-run''. These methods should not alter the scan results in
any way, but just use the information from the scanning process. A typical use
case would be to present the data in a nicer to use format than the standard
report, to use more external data sources or generate graphical representations
of data.

The post-run methods have the type \texttt{postrun} in the configuration, for
example:

\begin{verbatim}
[hexdump]
type        = postrun
module      = bat.generatehexdump
method      = generateHexdump
noscan      = text:xml:graphics:pdf:audio:video
envvars     = BAT_REPORTDIR=/tmp/images:BAT_IMAGE_MAXFILESIZE=100000000
description = Create hexdump output of files
enabled     = no
storetarget = reports
storedir    = /tmp/images
storetype   = -hexdump.gz
cleanup     = no
\end{verbatim}

\section{Scan configuration}

The analysis process is highly configurable: methods can be simply enabled and
disabled, based on need: some methods can run for quite a long time, which
might be undesirable at times. Configuration is done via a simple configuration
file in Windows INI format.

Most sections are specific to scanning methods, except two sections: a global
section and one section specific for the viewer tool.

\subsection{Global configuration}

The global configuration section is called \texttt{batconfig}. In this section
various global settings are defined which are described below. The section can
be identified in the configuration file by looking for this:

\begin{verbatim}
[batconfig]
\end{verbatim}

\subsubsection{\texttt{multiprocessing} and \texttt{processors}}

The \texttt{multiprocessing} configuration option determines whether or not
multiple CPUs (or cores) should be used during scanning. The default
configuration as shipped in the official BAT distribution is to use multiple
threads:

\begin{verbatim}
multiprocessing = yes
\end{verbatim}

If set to \texttt{yes} the program will start an extra process per CPU that is
available for parts of the program that can be run in parallel. In most cases
it is completely safe to use multiprocessing.

It might be desirable to not use all processors on a machine, for example if
there are multiple scans of BAT running at the same time, or if other tasks
need to run on the machine. It is possible to set the maximum amount of
processors to use with the \texttt{processors} option:

\begin{verbatim}
processors = 2
\end{verbatim}

\subsubsection{\texttt{writeoutputfile}}

Sometimes it is useful to let BAT just unpack, but not write an output file,
saving time packing data and writing it to disk. To prevent an output archive
being created set \texttt{writeoutputfile} to \texttt{no}. By default it will
be set to \texttt{yes}.

\subsubsection{\texttt{outputlite}}

Another setting in this section is \texttt{outputlite}:

\begin{verbatim}
outputlite = yes
\end{verbatim}

It defaults to \texttt{yes}. If set to \texttt{yes} the output archive will omit
a full copy of the unpacked data, significantly decreasing the size of the
output archive, but making it harder to do a ``post mortem'' on the unpacked
data (a new analysis should be run to get it again).

\subsubsection{\texttt{configdirectory}}

BAT allows configurations for scans to be split in different files and stored
in a separate directory. The directory with configurations can be set using the
\texttt{configdirectory} parameter:

\begin{verbatim}
configdirectory = /home/bat/configs/
\end{verbatim}

Important note: the configuration files \textit{have} to use the extension
\texttt{.conf}.

\subsubsection{\texttt{unpackdirectory}}

The unpacking directory where BAT will store its runtime state can be set using
\texttt{unpackdirectory}. By default this is \texttt{/tmp}. Each run of BAT
will create a new directory underneath this directory. In some cases it can be
wise to change it to a different location, with more storage (more and more
Linux distributions have \texttt{/tmp} mounted on a ramdisk) or less latency
(SSD). It can be use as follows:

\begin{verbatim}
unpackdirectory = /ssd/tmp
\end{verbatim}

\subsubsection{\texttt{temporary\_unpackdirectory}}

There is one setting to set the prefix for creating temporary files or
directories, namely \texttt{temporary\_unpackdirectory}. By default the
directory for creating temporary files and directories is \texttt{/tmp}. There
might be situations where the temporary directory might need to be changed, for
example for unpacking on a faster medium (ramdisk, SSD) than a normal harddisk.
It can be used as follows:

\begin{verbatim}
temporary_unpackdirectory = /ramdisk
\end{verbatim}

\subsubsection{\texttt{debug} and \texttt{debugphases}}

To assist in debugging and finding errors in scans of BAT there are two
settings: \texttt{debug} and \texttt{debugphases}. The setting \texttt{debug}
can be used to enable and disable debugging. If set multiprocessing will be
disabled and information about which file is scanned and which method is run
will be printed on standard error. If specified without \texttt{debugphases}
this will apply to all scan phases. The \texttt{debugphases} parameter can be
used to limit this behaviour to just one or a few phases. The other phases will
behave normally. For example, this will enable debugging, but just for the
leaf scans and aggregate scans:

\begin{verbatim}
debug = yes
debugphases = leaf:aggregate
\end{verbatim}

\subsubsection{\texttt{postgresql\_user}, \texttt{postgresql\_password}, \texttt{postgresql\_db}, \texttt{postgresql\_host} and \texttt{postgresql\_port}}

The PostgreSQL database used by BAT is configured in the global section. A few
variables have to be set to be able to connect with the database server, namely
the username, password and database name:

\begin{verbatim}
postgresql_user     = bat
postgresql_password = bat
postgresql_db       = bat
\end{verbatim}

Optionally a port and host can be set too if another port and/or host need to
be used:

\begin{verbatim}
postgresql_host     = 127.0.0.1
postgresql_port     = 5432
\end{verbatim}

Depending on the version of \texttt{python-psycopg2} it could be that
\texttt{postgresql\_host} and \texttt{postgresql\_port} both have to be
specified. For example on CentOS 6.x both parameters have to be set when using
a different port, even if the database resides on the local machine.

\subsubsection{\texttt{usedatabase}}

By default the database is enabled. There could be situations where it is
undesirable to use the database and it needs to be temporarily disabled. By
setting \texttt{usedatabase} to anything but \texttt{yes} the database will
be disabled:

\begin{verbatim}
usedatabase = no
\end{verbatim}

\subsubsection{\texttt{reporthash}}

If \texttt{reporthash} is set, then hashes in the ranking scan that come from
the BAT database will be converted from SHA256 (default) to the hash if
supported (currently MD5, SHA1 and CRC32 are supported) in the default BAT
database as created by the database scripts.

\begin{verbatim}
reporthash = sha256
\end{verbatim}

\subsubsection{\texttt{reportendofphase}}

If \texttt{reportendofphase} is set to \texttt{yes}, then BAT will write a
line with some statistics about when a scanning phase has ended on standard
output. This can be useful to track progress of BAT.

\begin{verbatim}
reportendofphase = yes
\end{verbatim}

\subsubsection{\texttt{packconfig} and \texttt{scrub}}

The output archive by default does not contain the configuration file that was
used during the scan. In some situations it is actually desirable to store the
configuration with the scan archive results, for example to debug an issue, or
to recreate results with the same configuration. For this the option
\texttt{packconfig} can be set to \texttt{yes}.

\begin{verbatim}
packconfig = yes
\end{verbatim}

Because the configuration file can contain confidential information (such as
database credentials) it is desirable to scrub this information from the
configuration file. For this the \texttt{scrub} setting can be used. Its value
should be a colon separated list of configuration options for which the value
in the configuration file (all occurances) should be replaced. For example, to
scrub the values of \texttt{postgresql\_user} and \texttt{postgresql\_password}
the following would be used:

\begin{verbatim}
scrub = postgresql_user:postgresql_password
\end{verbatim}

\subsubsection{\texttt{template}}

Some compression formats or file systems are stored anonymously without a name.
Examples are certain gzip-compressed files (like a ramdisk), or an LZMA-stream.

\begin{verbatim}
template = unpacked-by-bat-from-%s
\end{verbatim}

\subsubsection{\texttt{scansourcecode}}

The \texttt{scansourcecode} option can be used to check if a file that is
scanned can actually be found in the BAT database of source code files:

\begin{verbatim}
scansourcecode = yes
\end{verbatim}

The underlying rationale for this option is that various people have tried to
use BAT for source code scanning and did not get the results they expected. By
filtering out exact matches to the BAT database beforehand there are fewer
false positives.

\subsubsection{\texttt{dumpoffsets}}

During the BAT marker scans a dictionary with possible offsets for compressed
files and file systems is generated. Although most of these are discarded
during unpacking (as they are false positives) it could be useful to store this
data. By setting \texttt{dumpoffsets} to \texttt{yes} the offsets will be
stored as Python pickles in the \texttt{offsets} directory:

\begin{verbatim}
dumpoffsets = yes
\end{verbatim}

\subsubsection{\texttt{cleanup}}

If \texttt{cleanup} is set to \texttt{yes} the BAT working directory will be
removed after the scan has finished. By default the working directories are not
removed:

\begin{verbatim}
cleanup = yes
\end{verbatim}

\subsubsection{\texttt{compress}}

BAT outputs several result files. To save disk space these can be compressed
using gzip, at the expense of processing time. If not specified in the
configuration file \texttt{compress} will default to \texttt{no}.

\subsubsection{\texttt{packpickles}}

BAT internally uses Python pickles to (temporarily) store information on
disk. Since recent versions of BAT the preferred reporting and data exchange
format is JSON and pickles are no longer needed and don't need to be packed in
the output archive. If not specified in the configuration file
\texttt{packpickles} will default to \texttt{no}.

\subsubsection{\texttt{markersearchminimum}}

When a file is scanned for markers it is done in a single process. If many
files have to be scanned for markers at once this makes sense. However, if
there are multiple processors available and the top level file is big, then
it is a bit of a waste of time to not be able to use the extra processor
power. With \texttt{markersearchminimum} it is possible to set a minimum
size for a file to search for markers in the top level file in parallel. By
default top level files larger than 20 million bytes are processed in
allel.

\subsubsection{\texttt{tasktimeout}}

Several of the scanning phases in BAT use a task queue. Unfortunately it could
be that due to unknown bugs in BAT there are uncaught errors, which make it
seem like BAT hangs. For this the task queues have time outs. The default for
the timeout is 2592000 seconds (roughly one month). The task queue timeout can
be shortened by setting \texttt{tasktimeout} to a lower value:

\begin{verbatim}
tasktimeout = 2592000
\end{verbatim}

It should be noted that the value should not be set to 0, because otherwise
the queues will timeout immediately and BAT will barf.

\subsubsection{\texttt{tlshmaxsize}}

Beyond a certain size it no longer makes sense to compute TLSH checksums for
files. Using \texttt{tlshmaxsize} this limit can be set. By default it is set
to 52428800 bytes (50 MiB).

\subsubsection{Global environment variables}

Global environment variables are shared between scans. They can be overridden
by individual scans. For example to set the environment variable \texttt{FOO}
for all scans you would put something like this in the global configuration:

\begin{verbatim}
envvars     = FOO=/home/bat/bar
\end{verbatim}

To pass two or more environment variables use a semicolon:

\begin{verbatim}
envvars     = FOO=/home/bat/bar:XYZZY=1
\end{verbatim}

As a rule of thumb: settings that are shared between all scans should be set in
the global sections, while scan specific options should be in the scan
specific sections.

\subsection{Viewer configuration}

The other global section is \texttt{viewer}. This section is specific for the
graphical frontend and is not used in any other parts of BAT and might be moved
to a separate configuration file in a future version of BAT.

\subsection{Enabling and disabling scans}

The standard configuration file enables most of the scans and methods
implemented in BAT by default. Scans can be enabled and disabled by setting the
option \texttt{enabled} to \texttt{yes} and \texttt{no} respectively.

Another way to not run a scan is to outcomment the entry in the configuration
file (by starting the line with the \texttt{\#} character), or by removing the
section from the configuration file.

\subsection{Blacklisting and whitelisting scans}

Files can be explicitely blacklisted for scanning by using the \texttt{noscan}
configuration setting. The value of this parameter is a list of tags, separated
by colons:

\begin{verbatim}
noscan      = text:xml:graphics:pdf:audio:video
\end{verbatim}

Similarly files can be whitelisted by using the \texttt{scanonly} setting. Only
files that are tagged with any of the values in this list (if not empty) will
be scanned. If there is an overlapping value in \texttt{scanonly} and
\texttt{noscan} then the file will not be scanned.

\subsection{Passing environment variables}

All scans have an optional parameter \texttt{scanenv} defaulting to an empty
Python dictionary. In the configuration file a colon separated list of
name/value pairs can be specified using the keyword \texttt{envvars}. These will
then become available in the environment of the scan:

\begin{verbatim}
envvars     = BAT_REPORTDIR=/tmp/images:BAT_IMAGE_MAXFILESIZE=100000000
\end{verbatim}

If the environment of a scan needs to be adapted in the context of a single
file it is important to first make a copy of the environment or the
environment might be modified for the scan for all other files that are scanned.

\subsection{Scan names}

The name of the scan is used in various places, for example for storing results
or for determining scan conflicts. The \texttt{name} parameter can be used to
set the name for the scan. If no name is specified the name of the section of
the scan is used instead.

\begin{verbatim}
name = gzip
\end{verbatim}

\subsection{Scan conflicts}

Possibly scans can conflict with other scans in the same phase and they should
not be enabled at the same time. To indicate that a scan conflicts with others
the \texttt{conflicts} option can be set:

\begin{verbatim}
conflicts = gzip:bzip2
\end{verbatim}

If there is a conflict in the configuration BAT will refuse to run. Currently
BAT only looks at conflicts in the same unpacking phase and only for scans that
are enabled.

\subsection{Storing results}

Postrun scans and aggregate scans that output data, for example graphics files
or reports, can specify which files should be added to the output file. There
are three settings that should be set together:

\begin{verbatim}
storetarget = images
storedir    = /tmp/images
storetype   = -piechart.png:-version.png
\end{verbatim}

The \texttt{storetarget} setting specifies the relative directory inside the
output TAR archive. The \texttt{storedir} setting tells where to get the files
that need to be stored can be found (this should be where the postrun scan
or aggregate scan stores its results). The \texttt{storetype} setting is a colon
separated list of extensions/partial file names that the files should end in
(typically the rest of the filename is a SHA256 value).

The additional setting \texttt{cleanup} can be used to instruct BAT that the
files generated by this postrun scan or aggregate scan should be removed after
copying them into the result archive:

\begin{verbatim}
cleanup     = yes
\end{verbatim}

The \texttt{cleanup} setting should be set to \texttt{yes} unless the results
do not change in between subsequent runs of BAT.

Currently (BAT 36) if \texttt{cleanup} is set the files are written directly to
output directories. The values of these directories are hardcoded (and match
values that the GUI expects) but these will be replaced by the value of
\texttt{storetarget} in a later release.

\subsection{Running setup code}

For some scans it is necessary to run some setup code to ensure that certain
conditions are met, for example to see if database tables exist, or if locations
are readable/writeable. These checks only need to be run once. Based on the
result of the setup code the scan might be disabled if certain conditions are
not met.

There is a special hook for unpack scans, leaf scans and aggregate scans to run
setup code for the scan:

\begin{verbatim}
setup     = nameOfSetupMethod
\end{verbatim}

The result of the setup method is a tuple containing a boolean to indicate
wheter or not the scan should be run, and a (possibly adapted) environment.

The files \texttt{bat/identifier.py} and \texttt{bat/licenseversion.py} contain
very extensive examples of setup hooks.

\section{Analyser internals}

The analyser was written with extensibility in mind: new file systems or
variants of old ones tend to appear regularly (for example: there are at
least 5 or more versions of SquashFS with LZMA compression out there), and
sometimes it is needed to plug in a new unpacker for a file system or
compressed file type.

\subsection{Code organisation}

\texttt{bat-scan} is merely a frontend for the real scanner and only handle
the list of scans, the binary/binaries to scan and where to write the output
file(s).

The meaty bits of the analyser can be found in files in the \texttt{bat}
subdirectory (note that this directory currently contains more files than are
actually used by BAT at the moment):

\begin{itemize}
\item \texttt{batxor.py} contains experimental code to deal with files that
have been obfuscated with XOR.
\item \texttt{bruteforcescan.py} contains the main logic of the program: it
launches scans based on what is inside the binary and the scans that are
enabled, collects results from scans and writes results to an output file.
\item \texttt{busybox.py} and \texttt{busyboxversion.py} contain code to
extract useful information from a BusyBox binary, such as the version number.
\item \texttt{checks.py} contains various leaf scans, like scanning for certain
marker strings, or the presence of license texts and URLs of
forges/collaborative software development sites.
\item \texttt{ext2.py} implements some functionality needed for unpacking ext2
file systems.
\item \texttt{extractor.py} provides convenience functions that are used
throughout the code.
\item \texttt{file2package.py} has code to match names of files to names of
packages from popular distributions in a database.
\item \texttt{findduplicates.py} is used to find duplicate files in the scanned
archive.
\item \texttt{findlibs.py} and \texttt{interfaces.py} are for researching
dynamically linked ELF files in the archive.
\item \texttt{fixduplicates.py} is used to correct tagging of files that were
tagged incorrectly as duplicates, as they are the original, not the copy. For
now this is only for ELF files.
\item \texttt{fsmagic.py} contains identifiers of various file systems and
compressed files, like magic headers and offsets for which might need to be
corrected.
\item \texttt{fwunpack.py} includes most of the functionality for unpacking
compressed files and file systems.
\item \texttt{generatehexdump.py} and \texttt{images.py} generate textual and
graphical representations of the input files.
\item \texttt{generatereports.py}, \texttt{generateimages.py},
\texttt{guireport.py}, \texttt{generatejson.py} and \texttt{piecharts.py}
generate textual and graphical representations of results of the analysis.
\item \texttt{identifier.py} implements functionality to extract identifiers
(string constants, function names, method names, variable names, and so on)
from binary files and make them available for further analysis.
\item \texttt{javacheck.py} has code to parse Java class files.
\item \texttt{jffs2.py} has code specific to handling JFFS2 file systems.
\item \texttt{kernelanalysis.py} includes code to extract information from
Linux kernel images and Linux kernel modules.
\item \texttt{kernelsymbols.py} is used for generating dependency graphs for
Linux kernel modules and indicating any possible license issues of exported
symbols and declared licenses.
\item \texttt{licenseversion.py} gets version and licensing information for
uniquely identified strings and function names (and in the future variable names
too) from the database. It can optionally prune the result set to only include
relevant versions. It also contains code to aggregate results of Java class
files from a JAR file and assign results to the JAR file instead of the
individual class files.
\item \texttt{prerun.py} contains scans that are run in the pre-run phase for
correctly tagging files as early in the process as possible.
\item \texttt{prunefiles.py} can be used to remove files with a certain tag
from the scan results. This is useful for for example graphics files.
\item \texttt{renamefiles.py} is used for renaming files to use a more logical
name after more contextual information from the scan has become available. For
example: detect an initramfs in the Linux kernel and rename the temporary file
to \texttt{initramfs}.
\item \texttt{security.py} contains several security scans.
\item \texttt{unpackrpm.py} has code specifically for unpacking RPM archives.
\end{itemize}

\subsection{Pre-run methods}

Pre-run methods check and tag files, so the files can be ignored by later
methods and scans, reducing scanning time and preventing false positives. While
tagging is not exclusive to pre-run methods it is their main purpose.

\subsubsection{Writing a pre-run method}

Pre-run methods have a strict interface. Parameters are:

\begin{itemize}
\item \texttt{filename} is the absolute path of the file that needs to be tagged
\item \texttt{tempdir} is the (possibly) empty name of a directory where the
file is. This is currently unused and might be removed in the future.
\item \texttt{tags} is the set of tags that have already been defined for the
file.
\item \texttt{offsets} is the set of offsets that have been found for the file
\item \texttt{scanenv} is an optionally empty dictionary of environment variables
that can be used to pass extra information to the pre-run method.
\item \texttt{debug} is an environment variable that can be used to optionally
set the scan in debugging mode so it can print more information on standard
error. By default it is set to \texttt{False}.
\item \texttt{unpacktempdir} is the location of a directory for writing
temporary files. This value is optional and by default it is set to
\texttt{None}.
\end{itemize}

Return values are:

\begin{itemize}
\item a list containing tags
\end{itemize}

Example:

\begin{verbatim}
def prerunMethod(filename, tempdir=None, tags=[], offsets={},
                 scanenv={}, debug=False, unpacktempdir=None):
        newtags = []
        newtags.append('helloworld')
        return newtags
\end{verbatim}

\subsection{Unpackers}

Unpackers are responsible for recursively unpacking binaries until they
can't be unpacked any further.

\subsubsection{Writing an unpacker}

The unpackers have a strict interface:

\begin{verbatim}
def unpackScan(filename, tempdir=None, blacklist=[], offsets={},
               scanenv={}, debug=False):
        ## code goes here
\end{verbatim}

The last four parameters are optional, but in practice they are always passed
by the top level script.

\begin{itemize}
\item \texttt{tempdir} is the directory into which files and directories for
unpacking should be created. If it is \texttt{None} a new temporary directory
should be created.
\item \texttt{blacklist} is a list of byte ranges that should not be scanned.
If the current scan needs to blacklist a byte range it should add it to this
list after finishing a scan.
\item \texttt{offsets} is a dictionary containing a mapping from an identifier
to a list of offsets in the file where these identifiers can be found. This
list is filled by the scan \texttt{genericMarker} which always runs before
anything else.
\item \texttt{scanenv} is an optionally empty dictionary of environment variables
that can be used to pass extra information to the pre-run method.
\item \texttt{debug} is an environment variable that can be used to optionally
set the scan in debugging mode so it can print more information on standard
error. By default it is set to \texttt{False}.
\end{itemize}

Return values are:

\begin{itemize}
\item the name of a directory, containing files that were unpacked.
\item the blacklist, possibly appended with new values
\item a list of tags, in case any tags were added, or an empty list
\end{itemize}

Most scans have been split in two parts: one part is for searching the
identifiers, correctly setting up temporary directories and collecting results.
The other part is doing the actual unpacking of the data and verification.

The idea behind this split is that sometimes functionality is shared between
two scans. For example, \texttt{unpackCpio} is used by both
\texttt{searchUnpackCpio} and \texttt{unpackRPM}.

\subsubsection{Adding an identifier for a file system or compressed file}

Identifiers for new file systems and compressed files are, if available, added
to \texttt{fsmagic.py} in the directory \texttt{bat}. These identifiers will be
available in the \texttt{offsets} parameter that is passed to a scan, if any
were found.

Good sources to find identifiers are \texttt{/usr/share/magic}, documentation
for file systems or compressed files, or the output of \texttt{hexdump -C}.

\subsubsection{Blacklisting and priorities}

In BAT blacklists are used to prevent some scans from running on a particular
byte range, because other scans have already covered these bytes, or will cover
them.

The most obvious example is the \texttt{ext2} file system: in a normal setup
(no encryption) it is trivial to see the content of all the individual files
if an \texttt{ext2} file system image is opened. This is because this file
system is mostly a concatenation of the data, with some meta data associated
with the files in the file system.

If another compressed file is in the \texttt{ext2} file system it could be that
it will be picked up by BAT twice: once it will be detected inside the
\texttt{ext2} file system and once after the file system has been unpacked by
the \texttt{ext2} file system unpacker.

Other examples are:

\begin{itemize}
\item cpio (files are concatenated with a header and a trailer)
\item TAR (files are concatenated with some meta data)
\item RPM (files are in a compressed archive with some meta data)
\item ar and DEB
\item some flavours of cramfs
\item ubifs
\end{itemize}

To avoid duplicate scanning and false positives it is therefore necessary to
prevent other scans from running on the byte range already covered by one of
these files.

In BAT this is achieved by using blacklists. All unpackers have a parameter
called \texttt{blacklist} which is consulted every time a file is unpacked. If
a file system offset is in a blacklist the scan could use the next offset, or
skip scanning the entire file, depending on the scan.

The blacklist is set for every file individually and is initially empty. If a
scan is successful it adds a byte range to the blacklist. Subsequent scans
will skip the byte range added by the scan.

The scans are run in a particular order to make the best use of blacklists. The
order of scans is determined by the \texttt{priority} parameter in the
configuration file. The file systems and concatenated files mentioned above
have a higher priority and are scanned earlier than other scans that could also
give a match. It is not a fool proof system, but it seems to work well enough.

\subsection{Leaf scans}

After everything has been unpacked each file, including the files from which
other files were carved, will be scanned by the leaf scans.

\subsubsection{Writing a leaf scan}

The leaf scans have a simple interface. There are eight parameters passed to the
scan, namely the absolute path of the file, the tags of the file, a database
cursor and connection, an optional blacklist with byte ranges that should not
be scanned, an optional list of environment variables and an optional name of a
directory for writing temporary results. For example:

\begin{verbatim}
def leafScan(path, tags, cursor, conn, blacklist=[], scanenv={},
                debug=False, unpacktempdir=None):
        ## code goes here
\end{verbatim}

There are no restrictions on the return values of the leaf scan, except if
nothing could be found (in which case \texttt{None} is usd as return value).
The result value is a tuple with a list of tags as well as one of the following:

\begin{itemize}
\item \texttt{None} if nothing can be found
\item simple values (booleans, strings)
\item custom data structure. Code that processes this data should know about
its structure.
\end{itemize}

There is no restriction on the code that is run as part of the leaf scan and
basically anything can be done. In BAT there are for example checks that invoke
other external programs to discover dynamically linked libraries using
\texttt{readelf}, find the license of a kernel module using \texttt{modinfo}
or simple checks for the presence of strings in the binary that indicate the
use of certain software.

The simplest scans are the ones that search for hardcoded strings. These strings
are frequently found just in the package for which the check is written for. For
example, the following strings can often be found in copies of the
\texttt{iptables} binary and the related \texttt{libiptc} library:

\begin{verbatim}
markerStrings =
     [ 'iptables who? (do you need to insmod?)'
     , 'Will be implemented real soon.  I promise ;)'
     , 'can\'t initialize iptables table `%s\': \%s'
     ]
\end{verbatim}

Although searching for hardcoded strings is very fast, this method has some
drawbacks:

\begin{itemize}
\item a binary sometimes does not have these exact strings embedded
\item this method will only find the strings that are hardcoded and not any
other significant strings
\item if another package includes the string, it will be a false positive
\end{itemize}

The quick checks should therefore only be used as an indication that further
inspection of the binary is needed. A much better method is the ranking method
that is also available in BAT, but which requires a special setup with a
database.

\subsection{Aggregators}

Aggregators take all information from the entire scan process and possibly
modify results.

\subsubsection{Writing an aggregator}

Aggregators have a strict interface:

\begin{verbatim}
def aggregateexample(unpackreports, scantempdir, topleveldir, scanenv,
                     batcursors, batcons, debug=False, unpacktempdir=None)
\end{verbatim}

\begin{itemize}
\item \texttt{unpackreports} are the reports of the unpackers for all files
\item \texttt{scantempdir} is the location of the top level data directory of
the scan
\item \texttt{topleveldir} is the location of the top level directory of the
scan
\item \texttt{scanenv} is a dictionary of environment variables
\item \texttt{batcursors} is a list of PostgreSQL database cursors. If no
database is used this list will be empty.
\item \texttt{batcons} is a list of PostgreSQL database connections. If no
database is used this list will be empty.
\item \texttt{debug} is an environment variable that can be used to optionally
set the scan in debugging mode so it can print more information on standard
error. By default it is set to \texttt{False}.
\item \texttt{unpacktempdir} is the location of a directory for writing
temporary files. This value is optional and by default it is set to
\texttt{None}.
\end{itemize}

The aggregators should read any results of the leaf scans from the pickles
on disk.

If there is any result it should be returned as a dictionary with one key. It
will be assigned to the results of the top level element. Examples are: the
names of files which are duplicates in an archive or firmware.

\subsection{Post-run methods}

Post-run methods don't change the result of the whole scanning process, but
only use the data from the process. For example prettyprinting a fancy report
would be a typical post-run method.

\subsubsection{Writing a post-run method}

Post-run methods have a strict interface:

\begin{verbatim}
def postrunHelloWorld(filename, unpackreport, scantempdir, topleveldir,
                      scanenv, cursor, conn, debug=False):
        print "Hello World"
\end{verbatim}

\begin{itemize}
\item \texttt{filename} is the absolute path of the scanned file, after
unpacking.
\item \texttt{unpackreport} is the report of unpacking the file
\item \texttt{scantempdir} is the directory that contains the unpacked data
\item \texttt{topleveldir} is the top level directory containing the data
directory and the directory with the per file result pickles.
\item \texttt{scanenv} is a dictionary of environment variables
\item \texttt{cursor} is the database cursor (or \texttt{None} if there is no
database)
\item \texttt{conn} is the database connection (or \texttt{None} if there is no
database)
\item \texttt{debug} is an environment variable that can be used to optionally
set the scan in debugging mode so it can print more information on standard
error. By default it is set to \texttt{False}.
\end{itemize}

The post-run methods should read any results of the leaf scans from the pickles
stored on disk. Since the post-run methods don't change the result in any way,
but just have side effects there is no need to return anything. Any return value
will be ignored.

\section{Building binary packages of the Binary Analysis Tool}

If you want to install BAT through the package manager of your distribution you
might first need to generate packages for your distribution if none exist. For
BAT there is currently support to build packages for RPM-based systems and for
DEB-based systems.

\subsection{Building packages for RPM based systems from Git}

\subsubsection{Building \texttt{bat}}

Building the \texttt{bat} package is fairly straightforward.

\begin{enumerate}
\item Git clone the BAT repository from GitHub
\item \texttt{cd} to the directory \texttt{src}
\item run the command: \texttt{python setup.py bdist\_rpm}
\end{enumerate}

This will create an RPM file and an SRPM file. If you need to install BAT on
other versions of Fedora or on other RPM based distributions you can simply
rebuild the SRPM using:

\begin{verbatim}
rpmbuild --rebuild
\end{verbatim}

\subsubsection{Building \texttt{bat-extratools}}

Building packages for \texttt{bat-extratools} is unfortunately a bit more elaborate.

\begin{enumerate}
\item Git clone the \texttt{bat-extratools} repository from GitHub
\item change the name of \texttt{bat-extratools} directory to contain the version name of the
release (for example \texttt{bat-extratools-14.0}).
\item Make a \texttt{tar.gz} archive of the directory:

\begin{verbatim}
tar zcf bat-extratools-14.0.tar.gz bat-extratools-14.0
\end{verbatim}

\item run \texttt{rpmbuild} to create binary packages:

\begin{verbatim}
rpmbuild -ta bat-extratools-14.0.tar.gz
\end{verbatim}
\end{enumerate}

\subsection{Building packages for DEB based systems from releases}

Currently no rebuildable packages for DEB based systems are made for releases.

\subsection{Building packages for DEB based systems from Subversion}

\subsubsection{Building \texttt{bat}}

The Debian scripts were written according to the documentation for
\texttt{debhelper} found at \url{https://wiki.ubuntu.com/PackagingGuide/Python}.

Package building and testing is done on Ubuntu 14.04 LTS. Older versions of
Ubuntu are no longer supported and its use is discouraged. This is because
versions of Ubuntu older than 14.04 use a broken version of the PyDot package.

To build a \texttt{.deb} package do an export of the Subversion repository
first.  Change to the directory \texttt{src} and type:
\texttt{debuild -uc -us} to build the package. This assumes that you will have
the necessary packages installed to build the package (like \texttt{devscripts}
and \texttt{debhelper}).

The build process might complain about not being able to find the original
sources. In our experience it is safe to ignore this. The command will build a
\texttt{.deb} package which can be installed with \texttt{dpkg -i}.

\subsubsection{Building \texttt{bat-extratools}}

To build a .deb package clone the Git repository first. Change
to the correct directory (\texttt{bat-extratools} and type:
\texttt{debuild -uc -us} to build the package.

There are some dependencies that need to be installed beforehand, such as
\texttt{zlib1g-dev}, \texttt{liblzo2-dev} and \texttt{liblzma-dev} for building
\texttt{bat-extratools}. These dependencies are documented in the file
\texttt{debian/control} and \texttt{debuild} will warn if these packages are
missing.

\section{Binary Analysis Tool knowledgebase}

BAT comes with a mechanism to use a database backend. The default version of
BAT only unpacks file systems and compressed files and runs a few simple checks
on the leaf nodes of the unpacking process.

In the paper ``Finding Software License Violations Through Binary Code Clone
Detection'' by Hemel et. al. (ACM 978-1-4503-0574-7/11/05), presented at
the Mining Software Repositories 2011 conference, a method to use a database
with strings extracted from source code was described. This functionality is
available in the ranking module in the file \texttt{licenseversion.py}. This
code is enabled by default, but if no database is present it will not do
anything.

To give good results the database that is used needs to be populated with as
many packages as possible, from a cross cut of all of open source software, to
prevent bias towards certain packages: if you only would have BusyBox in your
database, everything would look like BusyBox.

A more detailed description about how to create the database can be found
in the BAT GitHub repository in the \textit{doc}

\subsection{Generating the package list}

The code and license extractor wants a description file of which packages to
process. This file is hardcoded to \texttt{LIST} relative to the directory that
contains all source archives. The reason there is a specific file is that some
packages do not follow a consistent naming scheme. By using this extra file we
can cleanup names and make sure that source code archives are recognized
correctly.

The file contains four values per line:

\begin{itemize}
\item name
\item version
\item archivename
\item origin (defaults to ``unknown'' if not specified)
\end{itemize}

separated by whitespace (spaces or tabs). An example would look like this:

\begin{verbatim}
amarok	2.3.2	amarok-2.3.2.tar.bz2	kde
\end{verbatim}

This line says that the package is \texttt{amarok}, the version number is
\texttt{2.3.2}, the filename is \texttt{amarok-2.3.2.tar.bz2} and the file
was downloaded from the KDE project.

There is a helper script (\texttt{generatelist.py}) to help generate the file.
It can be invoked as follows:

\begin{verbatim}
python generatelist.py -f /path/to/directory/with/sources -o origin
\end{verbatim}

The output is printed on standard output, so you want to redirect it to a file
called \texttt{LIST} (as expected by the string extraction script) and
optionally sorting it first:

\begin{verbatim}
python generatelist.py -f /path/to/directory/with/sources
    -o origin | sort > /path/to/directory/with/sources/LIST
\end{verbatim}

\texttt{generatelist.py} tries to determine the name of the package by
splitting the file name on the right on a \texttt{-} (dash)
character. This is not always done correctly because a package uses multiple
dashes, or because it does not contain a dash. In the latter case an error
will be printed on standard error, informing you that a file could not be
added to the list of packages and it should be added manually.

It is advised to manually inspect the file after generating it to ensure the
correctness of the package names. Packages can have been renamed for a number
of reasons:

\begin{itemize}
\item upstream projects decided to use a new name for archives (AbiWord
archives for example were renamed from \texttt{abi-\$VERSION.tar.gz} (used for
early versions) to \texttt{abiword-\$VERSION.tar.gz}).
\item a distribution has renamed packages to avoid clashes during installation
and allow different versions to be installed next to eachother.
\item a distribution has renamed a package. For example, Debian renamed
\texttt{httpd} to \texttt{apache2}.
\end{itemize}

In these cases you need to change the names of the packages, otherwise
different versions of the same package will be recorded in the database as
different packages, which will confuse the rating algorithm and cause it to
give suboptimal results.

Other helper scripts are \texttt{dumplist.py} which recreates a package list
file from a database, and \texttt{rewritelist.py} which takes two package list
files and outputs a new file with package names and versions rewritten for
filenames that occur in both files. These two scripts are useful if a database
needs to be regenerated, possibly with new packages.

\subsection{Creating the database}

The program to extract strings from sourcecode is \texttt{createdb.py}. It is
not part of the standard installation of BAT, but needs to be retrieved
separately from version control together with \texttt{generatelist.py}. This
will be changed at some point in the future.

It parses the file generated by \texttt{generatelist.py}, unpacks the files
(gzip compressed TAR, bzip2 compressed TAR, LZMA compressed TAR, XZ compressed
TAR and ZIP are currently supported) and scans each individual source code file
(written in C, C++, assembler, QML, C\#, Java, Scala, JSP, Groovy, PHP, Python,
Ruby and ActionScript) for string constants, methods, functions, variables and,
if enabled, licenses using Ninka and FOSSology and copyright information using
FOSSology and regular expressions lifted from FOSSology.

For the Linux kernel additional information is extracted about kernel functions
and variables, module information (author, license, parameters, and so on),
and kernel symbol information.

\subsection{License extraction and copyright information extraction}

The configuration for \texttt{createdb.py} has a few options. The most important
ones to consider are whether or not to also extract licenses and copyrights from
the source code files.  License extraction is done using the Ninka license
scanner and the Nomos license scanner from FOSSology. Copyright scanning is done
using the copyright scanner from FOSSology. These options are disabled
by default for a few reasons:

\begin{itemize}
\item extracting licenses and copyrights costs significantly more time
\item there are no packages for Fedora and Debian/Ubuntu for Ninka
\end{itemize}

If you want to enable license extraction, you will have to install Ninka first
and change one hardcoded path that points to the main Ninka script in
\texttt{createdb.py}. You will also have to install FOSSology
(for which packages are available for most distributions).

\subsection{Setting up PostgreSQL}

Setting up the PostgreSQL server itself is out of scope for this document. The
rest of this section should be considered as one potential way to set up a
PostgreSQL database. Any changes to the PostgreSQL installation should be
discussed with a local database administrator.

When starting from scratch (no existing database server) then the following
command can be used to initiate the database:

\begin{verbatim}
# postgresql-setup initdb
\end{verbatim}

\subsubsection{Authentication configuration}

The main authentication configuration file of PostgreSQL is called
\texttt{pg\_hba.conf}. Usually this file resides in the top level PostgreSQL
directory, for example \texttt{/var/lib/pgsql/data} (but this depends on the
local configuration). In this file various configuration options are set, such
as how clients can connect and how they should authenticate.

In a default installation of PostgreSQL local users connecting over a local
socket in the file system are implicitely trusted (this depends on the
distribution, some use \texttt{peer} or \texttt{ident} instead of
\texttt{trust}):

\begin{verbatim}
local   all      all       trust
\end{verbatim}

To prompt for the password (recommended) it should be changed in:

\begin{verbatim}
local   all      all       password
\end{verbatim}

Local users connecting over an IPv4 TCP/IP socket (over the network stack) are
also implicitely trusted:

\begin{verbatim}
host    all      all       127.0.0.1/32      trust
\end{verbatim}

and can be changed to:

\begin{verbatim}
host    all      all       127.0.0.1/32      password
\end{verbatim}

Something similar can be done for the local IPv6 connections.

To allow connections on a different port the main configuration file for
PostgreSQL (called \texttt{postgresql.conf}) should be adapted. By default the
server only listens on \texttt{localhost}, as defined by the
\texttt{listen\_address} configuration. To allow PostgreSQL to also listen on a
different interface this should be changed, for example:

\begin{verbatim}
listen_addresses = 'localhost,172.16.0.1'
\end{verbatim}

The authentication in \texttt{pg\_hba.conf} should also be changed:

\begin{verbatim}
host    all      all       172.16.0.1/16      password
\end{verbatim}

Note: in this case the network mask is \texttt{255.255.0.0}, but this could
of course be different.

\subsubsection{Creating the database and database user}

After setting up the PostgreSQL server the following commands can be used to
create a database and a user:

\begin{enumerate}
\item \texttt{create database bat;}
\item \texttt{create user bat with password 'bat';}
\item \texttt{grant all privileges on database bat to bat;}
\end{enumerate}

The database name, user name and password should correspond to the database
name, user name and password in the BAT configuration.

As a next step the database can be filled, either by loading an existing dump
file or by using the database creation scripts.

\subsection{Database design}

The database currently has 16 tables, 9 of which are Linux kernel specific.

\begin{itemize}
\item \texttt{processed}
\item \texttt{processed\_file}
\item \texttt{extracted\_string}
\item \texttt{extracted\_function}
\item \texttt{extracted\_name}
\item \texttt{kernel\_configuration}
\item \texttt{kernelmodule\_alias}
\item \texttt{kernelmodule\_author}
\item \texttt{kernelmodule\_description}
\item \texttt{kernelmodule\_firmware}
\item \texttt{kernelmodule\_license}
\item \texttt{kernelmodule\_parameter}
\item \texttt{kernelmodule\_parameter\_description}
\item \texttt{kernelmodule\_version}
\item \texttt{renames}
\item \texttt{hashconversion}
\item \texttt{extracted\_copyright}
\item \texttt{licenses}
\item \texttt{security}
\end{itemize}

\subsubsection{\texttt{processed} table}

This table is to keep track of which versions of which packages were scanned.
Its only purpose is to avoid scanning packages multiple times. It is not
actively used in the ranking code.

It has the following fields:

\begin{itemize}
\item \texttt{package}: name of the package
\item \texttt{version}: version of the package
\item \texttt{filename}: name of the archive
\item \texttt{origin}: site/origin where the archive was downloaded (optional)
\item \texttt{checksum}: SHA256 checksum of the archive
\item \texttt{downloadurl}: download URL of the source code package (optional)
\item \texttt{website}: the website of the project (optional)
\end{itemize}

\subsubsection{\texttt{processed\_file} table}
This table contains information about of individual source code files that were
scanned.

It has the following fields:

\begin{itemize}
\item \texttt{package}: name of the package the file is from (same as in
\texttt{processed})
\item \texttt{version}: version of the package the file is from (same as in
\texttt{processed})
\item \texttt{pathname}: relative path inside the source code archive
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{filename}: filename of the file, without path component
\item \texttt{thirdparty}: boolean
indicating if the file is an obvious copy of a file from another package.
\end{itemize}

\subsubsection{\texttt{extracted\_string} table}
This table stores the individual strings that were extracted from files and
that could possibly end up in binaries.

It has the following fields:

\begin{itemize}
\item \texttt{stringidentifier}: string constant that was extracted
\item \texttt{checksum}: SHA256 checksum of file the string constant was
extracted from
\item \texttt{language}: language the source code file was written in (mapped
to a language family, such as C or Java)
\item \texttt{linenumber}: line number where the string constant can be found
in the source code file (if determined using using \texttt{xgettext}) or $0$
(if determined using a regular expression).
\end{itemize}

\subsubsection{\texttt{extracted\_function} table}

In this table information about C functions and Java methods is stored.

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{functionname}: function name or method name that was extracted
\item \texttt{language}: language the source code file was written in (mapped
to a language family, such as C or Java)
\item \texttt{linenumber}: line number where the function/method can be found
in the source code file (if determined using using \texttt{xgettext}) or $0$
(if determined using a regular expression).
\end{itemize}

\subsubsection{\texttt{extracted\_name} table}
This table stores information of various names extracted from source code.
Included are variable names (C), field names (Java) and class names (Java) and
Linux kernel variable names.

It has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{name}: name of variable, field or class name that was extracted
\item \texttt{type}: type (field, variable, class name, etcetera)
\item \texttt{language}: language the source code file was written in (mapped
to a language family, such as C or Java)
\item \texttt{linenumber}: line number where the function/method can be found
in the source code file (if determined using using \texttt{xgettext}) or $0$
(if determined using a regular expression).
\end{itemize}

\subsubsection{\texttt{extracted\_copyright} table}
This table stores copyright information that was extracted from files by
FOSSology.

It has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{copyright}: copyright information that was extracted
\item \texttt{type}: type of information that was extracted, currently
\texttt{url}, \texttt{email} or \texttt{statement}
\item \texttt{offset}: byte offset in the file where the copyright statement
can be found
\end{itemize}

\subsubsection{\texttt{hashconversion} table}

The \texttt{hashconversion} table is used as a lookup table to translate
between different hashes and use these for checks or reporting. The
table has the following mandatory field:

\begin{itemize}
\item \texttt{sha256}: SHA256 checksum of the file
\end{itemize}

Any other hashes (limited to values that Python's hashlib supports, as well as
CRC32 and TLSH) listed in \texttt{extrahashes} in the database creation script
configuration file will be added as columns to this database. The database
creation scripts defaults to MD5, SHA1, CRC32 and TLSH.

\subsubsection{\texttt{kernel\_configuration} table}

The Makefiles in the Linux kernel configuration contain a lot of information
about which configuration includes which files. This information can be used to
reconstruct a possible kernel configuration that was used to create the Linux
binary image. The table has the following fields:

\begin{itemize}
\item \texttt{configstring}: configuration directive in Linux kernel
\item \texttt{filename}: filename/directory name to which the configuration
directive applies
\item \texttt{version}: Linux kernel version
\end{itemize}

\subsubsection{\texttt{kernelmodule\_alias} table}

This table is used to store information about Linux kernel module aliases. This
information is declared in the Linux kernel source code using the
\texttt{MODULE\_ALIAS} macro. The table has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{modulename}: name of the source code file
\item \texttt{alias}: contents of the \texttt{MODULE\_ALIAS} macro
\end{itemize}

\subsubsection{\texttt{kernelmodule\_author} table}

This table is used to store information about Linux kernel module author(s).
This information is declared in the Linux kernel source code using the
\texttt{MODULE\_AUTHOR} macro. The table has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{modulename}: name of the source code file
\item \texttt{author}: contents of the \texttt{MODULE\_AUTHOR} macro
\end{itemize}

\subsubsection{\texttt{kernelmodule\_description} table}

This table is used to store information about Linux kernel module descriptions.
This information is declared in the Linux kernel source code using the
\texttt{MODULE\_DESCRIPTION} macro. The table has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{modulename}: name of the source code file
\item \texttt{description}:
\end{itemize}

\subsubsection{\texttt{kernelmodule\_firmware} table}

This table is used to store information about Linux kernel module firmware.
This information is declared in the Linux kernel source code using the
\texttt{MODULE\_FIRMWARE} macro. The table has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{modulename}: name of the source code file
\item \texttt{firmware}: contents of the \texttt{MODULE\_FIRMWARE} macro
\end{itemize}

\subsubsection{\texttt{kernelmodule\_license} table}

This table is used to store information about Linux kernel module licenses.
This information is declared in the Linux kernel source code using the
\texttt{MODULE\_LICENSE} macro. The table has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{modulename}: name of the source code file
\item \texttt{license}: contents of the \texttt{MODULE\_LICENSE} macro
\end{itemize}

\subsubsection{\texttt{kernelmodule\_parameter} table}

This table is used to store information about Linux kernel module parameters.
This information is declared in the Linux kernel source code using the
\texttt{MODULE\_PARM} and \texttt{module\_param} macros, as well as variations
of the \texttt{module\_param} macro. These different notations were used for
different versions of the Linux kernel and both formats have been used in the
kernel at the same time. The table has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{modulename}: name of the source code file
\item \texttt{paramname}: name of the parameter
\item \texttt{paramtype}: type of the parameter, as specified in the source
code (various formats have been used)
\end{itemize}

\subsubsection{\texttt{kernelmodule\_parameter\_description} table}

This table is used to store information about Linux kernel module parameters
descriptions.  This information is declared in the Linux kernel source code
using the \texttt{MODULE\_PARM\_DESC} macro. The table has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{modulename}: name of the source code file
\item \texttt{paramname}: name of the parameter
\item \texttt{description}: descriptio of the parameter
\end{itemize}

\subsubsection{\texttt{kernelmodule\_version} table}

This table is used to store information about Linux kernel module versions.
This information is declared in the Linux kernel source code using the
\texttt{MODULE\_VERSION} macro. The table has the following fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{modulename}: name of the source code file
\item \texttt{version}: contents of the \texttt{MODULE\_VERSION} macro
\end{itemize}

\subsubsection{\texttt{licenses} table}

This table stores the licenses that were extracted from files using a source
code scanner, like Ninka or FOSSology. If a file has more than one licenses
there will be multiple rows for a file. It has these fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{license}: license as found by the scanner
\item \texttt{scanner}: scanner name. Currently only Ninka and FOSSology are
used in BAT, but is not limited to that: the scanner could also be a person
doing a manual review.
\item \texttt{version}: version of scanner. This is useful if there is for
example a bug in a scanner, or to compare results from various versions.
\end{itemize}

\subsubsection{\texttt{renames} table}

This is a lookup table to deal with packages that have been cloned or renamed
and should be treated as another package when scanning. Examples are packages
in Debian that have been renamed for trademark reasons (Firefox is called
Iceweasel), forks (KOffice versus Calligra), and so on.

\begin{itemize}
\item \texttt{originalname}: name the package was published under
\item \texttt{newname}: name that the package name should be translated to
\end{itemize}

The script \texttt{clonedbinit.py} in the \texttt{maintenance} directory
generates a minimal translation database.

\subsubsection{\texttt{security\_cert} table}

This table stores security information that was extracted from files. It has
these fields:

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{securitybug}: identifier for a security bug, for example
identifiers for the CERT secure coding standard.
\item \texttt{linenumber}: line number where the security bug can be found
\item \texttt{whitelist}: boolean value indicating whether or not the bug can
safely be ignored. The idea is that this can be set by security reviewers if
the security bug cannot be triggered to lower the amount of false positives.
\end{itemize}

\subsubsection{\texttt{security\_cve} table}

This table stores information about relations between paths and CVE numbers.

\begin{itemize}
\item \texttt{checksum}: SHA256 checksum of the file
\item \texttt{cve}: CVE identifier
\end{itemize}

\subsubsection{\texttt{security\_password} table}

This table stores information about relations between hashes and derived passwords.

\begin{itemize}
\item \texttt{hash}: hash value as found in password or shadow file
\item \texttt{password}: password found with a password cracker
\end{itemize}

\section{Identifier extraction and ranking scan}

As explained identifying binaries works in two phases: first identifiers are
extracted from the binaries, then the identifiers are processed by one or more
scans, for example the ranking scan.

Apart from making it possible to process the identifiers with various methods
there is another reason that the code is split in two parts and that is
performance: extracting identifiers is very quick and can be done in parallel
for many files. Computing a score can be quite expensive to do for certain
files (such as a Linux kernel image). Processing identifiers \textit{per file}
in parallel instead of processing files in parallel turns out to be much faster.
This is why the current ranking scan(s) are all aggregate scans and not leaf
scans.

\subsection{Configuring identifier extraction}

\begin{verbatim}
[identifier]
type        = leaf
module      = bat.identifier
method      = searchGeneric
envvars     = ramdisk:BAT_STRING_CUTOFF=5
noscan      = text:xml:graphics:pdf:compressed:
              resource:audio:video:mp4:vimswap:timezone:ico
description = Classify packages using advanced ranking mechanism
enabled     = yes
setup       = extractidentifiersetup
priority    = 1
\end{verbatim}

The parameter is:

\begin{itemize}
\item \texttt{BAT\_STRING\_CUTOFF} - this value is the mimimal length of the
string that is matched (default value is 5). If extracted strings are shorter
than this value they will be ignored. It is important to keep this parameter in
sync with the minimum length of strings in the database extract script.
\end{itemize}

\subsection{Configuring the ranking method}

The ranking method can be found in \texttt{bat/licenseversion.py}. The ranking
method looks up strings in the database, optionally aggregates results for
Java class files at the JAR level, determines versions and licenses while also
removing unlikely versions from the result set.

For the first part (determining which package a string belongs to) it uses
tables with caching information for string constants, function names, variable
names and so on. These caching tables contain a subset of information to vastly
speed up scanning by using pregenerated results to avoid expensive database
join operations. There is no script in the standard distribution of BAT to
create these caching tables, but the format has been described in the database
schema. For the second part (determining versions and licenses) other tables
containing the raw package data are used.

In the database the strings, averages, function names, variable names, etcetera
are split per language family (C, Java, C\#, and so on). The reason for this is
that strings/function names that are very significant in one programming
language family could be very generic in another programming language family
and vice versa. During scanning a guess will be made to see which language the
program was written in and the proper caching database will be queried.

Since there are relatively few binaries (at least on Linux) that combine
code from both languages the caching databases are split. This makes the
caching databases a lot smaller so they can easier fit into memory. There are
of course programs with language embeddeding and better support for these will
be added in the future.

An optional table in the database to deal with copied and renamed packages can
be generated with  \texttt{clonedbinit.py} in the \texttt{maintenance}
directory. If this table is populated the ranking scan will use information
from this table to rewrite package names. This is useful if a package was
renamed for a reason and different packages should be treated as if they were
a single package. Examples are Ethereal that had to be renamed to Wireshark,
or KOffice that was forked into Calligra, after which development on KOffice
effectively stopped and everyone moved to Calligra.

If \texttt{BAT\_RANKING\_LICENSE} is not set to 1 no license information will
be extracted. If \texttt{BAT\_RANKING\_VERSION} is not set to 1 no version
information will be extracted. If \texttt{BAT\_RANKING\_LICENSE} is set to 1 it
automatically sets \texttt{BAT\_RANKING\_VERSION} to 1 as well.

The parameter \texttt{USE\_SOURCE\_ORDER} can be used to tell the matching
algorithm to assume that identifiers in the binary code are similar as in the
source code and that the compiler has not reordered these. As compilers often
keep the order this assigns more strings to packages. As soon as compilers
start reordering identifiers this method will not work. The default setting
is to not use the order of identifiers.

The parameter \texttt{BAT\_STRING\_CUTOFF} indiciates the mimimal length of the
string that is matched (default value is 5). If extracted strings are shorter
than this value they will be ignored. It is important to keep this parameter in
sync with the minimum length of strings in the database extract script.

Results of Java class files are aggregated per JAR where the class files were
found in. If the parameter \texttt{AGGREGATE\_CLEAN} is set to 1 the class
files will be removed from the result set after aggregating the results. By
default class files will not be removed.

The parameters \texttt{BAT\_KEEP\_VERSIONS}, \texttt{BAT\_MINIMUM\_UNIQUE} and
\texttt{BAT\_KEEP\_MAXIMUM\_PERCENTAGE} are used to tell the pruning methods
how many versions to keep, how many unique strings minimally should be found,
and so on.

\subsubsection{Interpreting the results}

There are two ways to interpret the results. The recommended way is to load the
result file into the graphical user interface.

If there are any matches the report contains the following:

\begin{itemize}
\item number of lines that were extracted from the binary
\item number of lines that could be matched exactly with an entry in the
database (unique matches)
\item number of lines that were assigned to a package (assigned matches)
\item number of lines which could not be matched (unmatched lines)
\end{itemize}

Per package the following is reported:

\begin{itemize}
\item name of the package
\item all unique matches (strings that can only be found in this package)
\item relative ranking
\item percentage of the total score
\end{itemize}

Finally piecharts are also generated providing a visual representation of these results.

\section{BusyBox script internals}

The BusyBox processing scripts look simple, but behind the internals are a bit
hairy. Especially extracting the correct configuration is not trivial.

\subsection{Detecting BusyBox}

Detecting if a binary is indeed BusyBox is trivial, since in a BusyBox binary
there are almost always clear indication strings if BusyBox is used (unless
they it was specifically altered to hide the use of BusyBox).

A significant set of strings to look for is:

\begin{verbatim}
        BusyBox is a multi-call binary that combines many common Unix
        utilities into a single executable.  Most people will create a
        link to busybox for each function they wish to use and BusyBox
        will act like whatever it was invoked as!
\end{verbatim}

Another clear indicator is a BusyBox version string, for example:

\begin{verbatim}
BusyBox v1.15.2 (2009-12-03 00:14:42 CET)
\end{verbatim}

As an exception a BusyBox binary configured to include just a single applet
will not contain contain the marker strings, or the BusyBox version string. In
such a case a different detection mechanism will have to be used, for example
the ranking code as used by \texttt{bat-scan}, although this will only be
necessary in a very small percentage of cases, since the vast majority of
BusyBox instances include more than one applet.

\subsection{BusyBox version strings}

The BusyBox version strings have remained fairly consistent over the years:

\begin{verbatim}
BusyBox v1.00-rc2 (2006.09.14-03:08+0000) multi-call binary
BusyBox v1.1.3 (2009.09.11-12:49+0000) multi-call binary
BusyBox v1.15.2 (2009-12-03 00:14:42 CET)
\end{verbatim}

The time stamps in the version string are irrelevant, since they are generated
during build time and are not hardcoded in the source code.

Extracting version information from the BusyBox binary is not difficult.
Using regular expression it is possible to look for \texttt{BusyBox v} which
indicates the start of a BusyBox version string. The version number can be
found immediately following this substring until \texttt{ (} (including
leading space) is found.

Apart from reporting, the BusyBox version number is also used for other
things, such as determining the right configuration format and accessing a
knowledgebase of known applet names extracted from the standard BusyBox
releases from \url{busybox.net}.

\subsection{BusyBox configuration format}

During the compilation of BusyBox a configuration file is used to determine
which functionality will be included in the binary. The format of this
configuration file has changed a few times over the years. Early versions used
a simple header format file, with GNU C/C++ style defines. Later versions,
starting 1.00pre1, moved to Kbuild, the same configuration system as used by
for example the Linux kernel or OpenWrt. This format is still in use today
(BusyBox 1.20.0 being the latest version at the time of writing).

Each configuration directive determines whether or not a certain piece of
source code will be compiled and up in the BusyBox binary. This source code can
either be a full applet, or just a piece of functionality that merely extends
an existing applet.

\subsection{Extracting a configuration from a BusyBox binary}

Extracting the BusyBox configuration from a binary is not entirely trivial.
There are a few methods which can be used:

\begin{enumerate}
\item run \texttt{busybox} (on a device, or inside a sandbox) and see what
functionality is reported. This is probably the most accurate method, but also
the hardest, since it requires access to a device, or a sandbox that has been
properly set up, with all the right dependencies, and so on.

When running \texttt{busybox} without any arguments, or with the \texttt{--help}
parameter it will output a list of functions that are defined inside the
binary:

\begin{verbatim}
Currently defined functions:
        ar, cal, cpio, dpkg, dpkg-deb, gunzip, zcat
\end{verbatim}

These can be mapped to a configuration, using information extracted from
BusyBox source code about which applets map to which configuration option.
\item extract the configuration from the binary by searching for known applet
names in the firmware. The end result is the same as a previous step, but
possibly with less accuracy in some cases but it is the only feasible solution
if you only have a binary.
\end{enumerate}

The BusyBox binary has a string embedded for every applet that is included.
This is the string that is printed out if \texttt{--help} is given as a
parameter to an invocation of \texttt{busybox}.

Using information about the configuration extracted from BusyBox source code
these strings can be mapped to a configuration directive and a possible
configuration can be reconstructed.

Depending on how the binary was compiled this can be trivial, or quite hard.

\subsubsection{BusyBox linked with uClibc}

In binaries that link against uClibc (a particular C library) the name of the
main function of the applet is sometimes (but not always) included in the
\texttt{busybox} binary as follows (a good way is to run \texttt{strings} on
the binary and look at the output).

\begin{verbatim}
wget_main
\end{verbatim}

This string maps to the name of the main function for the \texttt{wget} applet
(\texttt{networking/wget.c}):

\begin{verbatim}
int wget_main(int argc, char **argv) MAIN_EXTERNALLY_VISIBLE;
\end{verbatim}

The BusyBox authors are pretty strict in their naming and usually have a
configuration directive in the a specific format
(\texttt{CONFIG-\$appletname}) in the Makefile, like:

\begin{verbatim}
lib-$(CONFIG_WGET)         += wget.o
\end{verbatim}

(example taken from \texttt{networking/Kbuild} in BusyBox 1.15.2). There are
cases where the format could be slightly different.

\subsubsection{BusyBox linked with glibc \& uClibc exceptions}

Sometimes the method described in the previous section does not work for
binaries that are linked with uClibc. It also does not work with binaries
compiled with glibc.

If the binary is unstripped and the binary still contains symbol information
it is possible to extract the right information using \texttt{readelf} (part
of GNU binutils) in a similar fashion as the earlier described method.

In case there is no information available it is still possible to search inside
the binary for the applet names. Because most instances of BusyBox that are
installed on devices have not been modified the list of applets in the stock
version of BusyBox serves as an excellent starting point.

The list as printed by \texttt{busybox} if the \texttt{--help} parameter is
given is embedded in the binary. The applet names are alphabetically sorted
and separated by NUL characters.

By searching for this list and splitting it accordingly it is possible to get
the list of all applets that are defined. The only caveats are that a new
applet that was added appears alphabetically before any of the applets that
can be recognized using a list of applet names extracted from the source code,
or it appears alphabetically after the last one that can be recognized.

\subsection{Pretty printing a BusyBox configuration}

Pretty printing a BusyBox configuration is fairly straightforward, but there
are a few cases where it is hard to make a good guess:

\begin{enumerate}
\item aliases
\item functionality that is added to an applet, depending on a configuration
directive
\item applets that use non-standard configuration names (like
\texttt{CONFIG\_APP\_UDHCPD} instead of \texttt{CONFIG\_UDHCPD} in some
versions of BusyBox)
\item features
\end{enumerate}

For some applets aliases are installed by default as symlinks. These aliases
are recorded in the binary, but there is no separate applet for it. In the
BusyBox sources (1.15.2, others might be different) these are defined as:

\begin{verbatim}
IF_CRYPTPW(APPLET_ODDNAME(mkpasswd, cryptpw, _BB_DIR_USR_BIN,
    _BB_SUID_DROP, mkpasswd))
\end{verbatim}

So if the \texttt{cryptw} tool is built, an additional symlink called
\texttt{mkpasswd} is added during installation.

If extra functionality is added to an applet in BusyBox it is defined in the
source code by macros like the following:

\begin{verbatim}
IF_SHA256SUM(APPLET_ODDNAME(sha256sum, md5_sha1_sum, _BB_DIR_USR_BIN,
    _BB_SUID_DROP, sha256sum))
IF_SHA512SUM(APPLET_ODDNAME(sha512sum, md5_sha1_sum, _BB_DIR_USR_BIN,
    _BB_SUID_DROP, sha512sum))
\end{verbatim}

The above configuration tells to add extra symlinks for \texttt{sha256sum} and
\texttt{sha512sum} if BusyBox is configured for suppport for the SHA256 and
SHA512 algorithms. The applet that implements this functionality is
\texttt{md5\_sha1\_sum}.

Non-standard configuration names can be fixed by using a translation table that
translates to the non-standard name. The current code has a translation table
for BusyBox 1.15 and higher.

Detecting features is really hard to do in a generic way. In most cases it will
even be impossible, because there are no clear markers (strings, applet names)
in the binary that indicate that a certain feature is enabled. In cases there
are clear marker strings these would still need to be linked to specific
features. One possibility would be to parse the BusyBox sources and link
strings to features, for example (from BusyBox 1.15.3,
\texttt{editors/diff.c}):

\begin{verbatim}
#if ENABLE_FEATURE_DIFF_DIR
       diffdir(f1, f2);
       return exit_status;
#else
       bb_error_msg_and_die("no support for directory comparison");
#endif
\end{verbatim}

The string \texttt{"no support for directory comparison"} only appears if the
feature \texttt{ENABLE\_FEATURE\_DIFF\_DIR} is not enabled.

Implementing this will be a lot of work and it will likely not be very useful.

\subsection{Using BusyBox configurations}

By referencing with information extracted from the standard BusyBox sourcecode
it is possible to get a far more accurate configuration, because it is known
which applets use which configuration, unless:

\begin{itemize}
\item new applets were added to BusyBox
\item applets use old names, but contain different code
\end{itemize}

The names of applets that are defined in BusyBox serve as a very good starting
point. How these are recorded in the sources has changed a few times and
depends on the version of BusyBox. The tool \texttt{appletname-extractor.py}
can extract these from the BusyBox sources and store them for later reference
as a simple lookup table in Python pickle format.

Names of applets per version breakdown:

\begin{itemize}
\item 1.15.x and later: \texttt{include/applets.h} or
\texttt{include/applets.src.h} IF syntax
\item 1.1.1-1.14.x: \texttt{include/applets.h} USE syntax
\item 1.00-1.1.0: \texttt{include/applets.h} (different syntax)
\item 0.60.5 and earlier: \texttt{applets.h}, like 1.00-1.1.0 but with a
slightly different syntax
\end{itemize}

In one particular version of BusyBox (namely 1.1.0) there is a mix of three
different syntaxes: (0.60.5, 1.00 and another) for a few applets
(\texttt{runlevel}, \texttt{watchdog}, \texttt{tr}).

There are also a few applets in 1.1.0 which seem to be a bit harder to detect:
\texttt{busybox}, \texttt{mkfs.ext3}, \texttt{e3fsck} and \texttt{[[}. These
can easily be added by hand, since there are just four of them.

Another issue that is currently unresolved is that not all the shells are
correctly recognized.

\subsection{Extracting configurations from BusyBox sourcecode}

The \texttt{busybox.py} script makes use of a table that maps applet names to
configuration directives. These tables are stored in a Python pickle and read
by \texttt{busybox.py} upon startup. To generate these pickle files the
\texttt{appletname-extractor.py} should be used. In the standard distribution
for BAT the configurations for most versions of BusyBox are shipped.

The applet names are extracted from a file called \texttt{applets.h} or
\texttt{applets.src.h}.

\begin{verbatim}
python appletname-extractor.py -a /path/to/applets.h -n $VERSION
\end{verbatim}

The configuration will be written to a file \texttt{\$VERSION-config} and
should be moved into the directory containing the other configurations.

\section{Linux kernel identifier extraction}

The \texttt{createdb.py} program processes Linux kernel source code files in a
slightly different way than normal source code files. There is a lot of
interesting information that can be extracted from the Linux kernel sources,
as well as the binary.

There are a few challenges when working with Linux kernel source code and Linux
kernel binaries. First of all there are many different variants in use and many
vendors have their own slightly modified version, with extra drivers, or bug
fixes from later versions, or bug fixes that might not yet have been applied to
the version on \url{kernel.org}.

Second is that in the Linux kernel binary string constants, function names,
symbols, module parameters, and so on, are intertwined and some steps need to
be taken to correctly split these to avoid false positives (there are other
packages where kernel function names, module parameters, symbols, and so on,
are valid string constants).

\subsection{Extracting visible strings from the Linux kernel binary}

If a kernel is an ELF binary (sometimes) the relevant sections of the
binary can be read using \texttt{readelf}. Otherwise \texttt{strings} can be
run on the binary. This method will return more strings than if using
\texttt{readelf}, but the extra strings are mostly extra cruft that have a low
chance of matching.

\subsection{Extracting visible strings from a Linux kernel module}

If a kernel module is an ELF binary (most cases) the relevant sections of the
binary can be read using \texttt{readelf}. Otherwise \texttt{strings} can be
run on the binary. This method will return more strings than if using
\texttt{readelf}, but the extra strings are mostly extra cruft that have a low
chance of matching.

\subsection{Extracting strings from the Linux kernel sources}

The Linux kernel is full of strings that can end up in a binary. Some
programmers have defined macros just specific to their part of the kernel for
ease of use (often a wrapper around \texttt{printk}, other programmers use
more standard mechanisms like \texttt{printk}. Most strings can be extracted
from the Linux kernel using \texttt{xgettext}. A minority of strings needs to
be extracted using a custom regular expression.

The following two cases are worth a closer look:

\subsubsection{\texttt{EXPORT\_SYMBOL} and \texttt{EXPORT\_SYMBOL\_GPL}}

The symbols defined in the \texttt{EXPORT\_SYMBOL} and
\texttt{EXPORT\_SYMBOL\_GPL} macros end up in the kernel image. The
\texttt{EXPORT\_SYMBOL\_GPL} symbol could be interesting for licensing
reporting as well, since anything that uses this symbol should be released
under the GPLv2. This is a topic for future research.

\subsubsection{\texttt{module\_param}}

The names of parameters for kernel modules can end up in the kernel, or in the
kernel module itself. The names of these parameters are typically prefixed
with the name of the module (which is often, but not always) and a dot, but
without the extension of the file. In cases where the module name does not
match the name of the file it was defined in extra information from the
build system needs to be added to determine the right string.

The code for this is in the function \texttt{\_\_init param\_sysfs\_builtin} in
\texttt{kernel/params.c}.

Module names are extracted from the kernel Makefiles and stored in the database
together with module information (author, license, description, parameters,
and so on).

\subsection{Forward porting and back porting}

There are some strings we scan for which might not be present in certain
versions, because they were removed, or not yet included in the mainline
kernel. A good example is devfs. This subsystem was removed in Linux kernel
2.6.17, but it is not safe to assume that this was done for every 2.6.17 (or
later) kernel that is out in the wild, since some vendors might have kept it
and ported it to newer versions (forward porting). Similarly code from newer
kernels might have been included in older versions
(backporting).

\subsection{Corner cases}

Sometimes a \texttt{\#define} or some configuration directive causes that our 
string matching method will not work, because the string is prepended with
extra characters.

An example from \texttt{arch/arm/mach-sa1100/dma.c} from kernel 2.6.32.9:

\begin{verbatim}
#undef DEBUG
#ifdef DEBUG
#define DPRINTK( s, arg... )  printk( "dma<%p>: " s, regs , ##arg )
#else
#define DPRINTK( x... )
#endif
\end{verbatim}

Other examples include \texttt{pr\_debug}, \texttt{DBG}, \texttt{DPRINTK} and
\texttt{pr\_info}.

To work around this there are two ways:

\begin{enumerate}
\item do substring matches
\item parse the source code and record where extra code is being added as in
the example above and only do substring matches in a small number of cases.
\end{enumerate}

Substring matching is expensive and since it only happens in a minority of
cases the second method, although not trivial to implement, would be easier.
This is future work.

\section{Binary Analysis Tool performance tips}

This section describes a few methods to increase performance of the Binary
Analysis Tool, plus describe drawbacks of methods named. The standard
configuration of BAT tries to be sensible, with a trade off between performance
and completeness. In some cases there is quite a bit of performance to be
gained by simply tweaking the configuration.

\subsection{Choose the right hardware}

BAT will benefit a lot from fast disk, enough memory and multiple cores. Many
of the scans in BAT can be run in parallel and will scale very well (until
of course disk I/O limits are reached). Invest in SSD to reduce disk I/O and
more cores instead of a faster CPU. Enough memory will prevent swapping which
just kills performance, especially because the ranking scan in BAT can be very
I/O intensive.

\subsection{Use \texttt{outputlite}}

Using the default configuration the original unpacked data is not included into
the result archive.

There are situations where it makes sense to include the data into the result
archive, for example to make it easier to do a ``post mortem'' after a scan.
The original data can take up a lot of space, since every original file, plus
everything that might have been extracted from that file, will be included,
which leads to large archives and long associated packing time.

It also has performance impact on the BAT viewer, which needs to unpack some
data from the archive. The smaller the archive is, the faster unpacking is.

If the original data and the unpacked data is not relevant, then setting the
option \texttt{outputlite} to \texttt{yes} in the section \texttt{[batconfig]}
is highly recommended:

\begin{verbatim}
outputlite = yes
\end{verbatim}

\subsection{Use \texttt{AGGREGATE\_CLEAN} when scanning Java JAR files}

If Java JAR files are scanned then pictures and reports will be generated for
each of the individual \texttt{.class} files. If only the results of the JAR
file are needed, then setting \texttt{AGGREGATE\_CLEAN} to \texttt{1} will
prevent pictures and reports to be generated for the individual \texttt{.class}
files, which can save quite some processing time and help declutter the
interface as well.

Of course, not generating the pictures for individual \texttt{.class} files
means that some detail might be lost, especially if there are \texttt{.class}
files that contain some unexpected results.

\subsection{Disable \texttt{tmp} on tmpfs}

Some Linux distributions (most notably Fedora 18 and later) store the
\texttt{/tmp} file system on tmpfs. This means that part of the system memory
is used for the \texttt{/tmp} file system. By default on Fedora it is
set to 50\% of the system's memory. This could influence BAT in two ways:

\begin{enumerate}
\item less memory available for processing
\item BAT unpacks to \texttt{/tmp} by default, unless configured differently. If
the unpack results grow big enough (which is fairly easy with big firmwares) it
could fill up the partition. However, there are some external tools that will
write temporary results to \texttt{/tmp}.
\end{enumerate}

There are various solutions, apart from adding more memory to the machine:

\begin{itemize}
\item configure BAT to use another path than \texttt{/tmp} for unpacking and
storing results and configure some scans in BAT to use \texttt{/tmp} or a
different ramdisk (recommended)
\item disable \texttt{tmp} on tmpfs (not recommended)
\end{itemize}

\subsection{Use tmpfs for writing temporary results}

A few scans can use tmpfs or a ramdisk to write temporary results. The scans
that can benefit from this are LZMA unpacking, ranking (temporary results of
DEX and ODEX unpacking), compress unpacking, JFFS2 unpacking and TAR unpacking.

The parameter \texttt{temporary\_unpackdirectory} in the global configuration
can be used to set this location.

\section{Description for scans using the database}

\subsection{\texttt{file2package}}

The \texttt{file2package} leaf scan uses the table \texttt{file} that contains
information with checksums of file names found in standard Linux distributions.
This table can be populated using the scripts
\texttt{createfiledatabasedebian.py} and \texttt{createfiledatabasefedora.py}
which can be found in the subdirectory \texttt{maintenance} in the BAT source
tree and are for Debian and Fedora respectively.

\section{Parameter description for default scans}

This section describes the default parameters for several of the scans as
shipped in BAT, if not described earlier in this document. These parameters are
passed to the scans as part of the environment and are defined in the
\texttt{envvars} setting in the configuration file.

\subsection{\texttt{compress}}

The \texttt{COMPRESS\_MINIMUM\_SIZE} parameter instructs the scan to ignore
output files that are \texttt{COMPRESS\_MINIMUM\_SIZE} bytes in size or less.
This parameter was introduced because false positives in compress unpacking are
very common on Debian and Ubuntu, often leading to small sized files that
contain no useful data and which could interfere with scanning.

The scan can also use a different directory for unpacking temporary files.
The location is set in the global configuration using the parameter
\texttt{temporary\_unpackdirectory}. By setting this to for example SSD or
ramdisk it can help avoid disk I/O on a slower disk and speed up scanning.

\subsection{\texttt{jffs2}}

The scan can use a different directory for unpacking temporary files.
The location is set in the global configuration using the parameter
\texttt{temporary\_unpackdirectory}. By setting this to for example SSD or
ramdisk it can help avoid disk I/O on a slower disk and speed up scanning.

\subsection{\texttt{lzma}}

The \texttt{lzma} unpack scan has one parameter: \texttt{LZMA\_MINIMUM\_SIZE}.

The \texttt{LZMA\_MINIMUM\_SIZE} parameter instructs the scan to ignore output
files that are \texttt{LZMA\_MINIMUM\_SIZE} bytes in size or less. This
parameter was introduced because false positives in LZMA unpacking are very
common, often leading to small sized files that contain no useful data.

By default \texttt{LZMA\_MINIMUM\_SIZE} is set to 10 bytes, but this is a very
conservative setting and can likely be set higher safely.

The scan can also use a different directory for unpacking temporary files.
The location is set in the global configuration using the parameter
\texttt{temporary\_unpackdirectory}. By setting this to for example SSD or
ramdisk it can help avoid disk I/O on a slower disk and speed up scanning.

\subsection{\texttt{tar}}

The scan can use a different directory for unpacking temporary files.
The location is set in the global configuration using the parameter
\texttt{temporary\_unpackdirectory}. By setting this to for example SSD or
ramdisk it can help avoid disk I/O on a slower disk and speed up scanning.

\subsection{\texttt{xor}}

The \texttt{XOR\_MINIMUM} parameter is used to set the minimum amount of
occurences of a key that have to be present in the file before XOR unpacking
is done. This is to reduce false positives.

\subsection{\texttt{zip}}

The \texttt{ZIP\_MEMORY\_CUTOFF} parameter is used to set the maximum size
of ZIP data that should be read into memory. If the ZIP data is larger it will
be carved out from the larger file using \texttt{dd}. If not set a value of 50
million bytes will be used.

\subsection{\texttt{findlibs}}

For the \texttt{findlibs} aggregate scan the \texttt{ELF\_SVG} parameter can be
set to \texttt{1} to output the graphs in SVG format.

\subsection{\texttt{findsymbols}}

For the \texttt{findsymbols} aggregate scan the \texttt{KERNELSYMBOL\_SVG}
parameter can be set to \texttt{1} to output the graphs in SVG format. The
\texttt{KERNELSYMBOL\_CSV} parameter can be set to output a spreadsheet in
Excel-format.

\subsection{\texttt{generateimages}}

The \texttt{generateimages} postrun scan has five optional parameters:
\texttt{AGGREGATE\_IMAGE\_SYMLINK}, \texttt{BAT\_IMAGEDIR}, \texttt{BAT\_PICKLEDIR}, \texttt{MAXIMUM\_PERCENTAGE}
\texttt{MINIMUM\_PERCENTAGE}

\subsection{\texttt{identifier}}

The scan can use a different directory for unpacking temporary files.
The location is set in the global configuration using the parameter
\texttt{temporary\_unpackdirectory}. By setting this to for example SSD or
ramdisk it can help avoid disk I/O on a slower disk and speed up scanning.

\subsection{\texttt{licenseversion}}

The \texttt{licenseversion} aggregate scan has a few parameters that can
influence performance. One of them is \texttt{AGGREGATE\_CLEAN}. This parameter
instructs the scan to remove results for individual Java class files from the
result set after aggregating results at the JAR level. Java class files that
are not unpacked from a JAR file are not removed from the result set. By
default this parameter is set to 0 which means that results for Java class
files are not removed from the result set.

\subsection{\texttt{prunefiles}}

The \texttt{prunefiles} aggregate scan has two parameters: \texttt{PRUNE\_TAGS}
and \texttt{PRUNE\_FILEREPORT\_CLEAN}. The \texttt{PRUNE\_TAGS} parameter
contains a comma-separated list of tags that should be ignored and removed from
the scan results. The \texttt{PRUNE\_FILEREPORT\_CLEAN} parameter can be set to
indicate whether or not the result pickles for the pruned files should also be
removed from disk. Example:

\begin{verbatim}
PRUNE_TAGS=png,gif:PRUNE_FILEREPORT_CLEAN=1
\end{verbatim}

\subsection{\texttt{hexdump} and \texttt{images}}

The \texttt{hexdump} and \texttt{images} scans (disabled by default) have two
parameters. The \texttt{BAT\_IMAGE\_MAXFILESIZE} parameter is set to specify
the maximum size of a file for which a result is generated. Since output from
this scan can be extremely large, and the results are not very interesting for
large files it is strongly advised to cap this value.

\section{Default ordering of scans in BAT}

BAT comes with a default configuration file. In this file an order for running
the scans is specified, using the \texttt{priority} field: the higher the
priority, the earlier the scan is run in the process. In this section the
rationale behind this ordering is explained.

The order for pre-run scans, leaf scans, unpack scans and aggregate scans is
described below. Since postrun scans do not change the result files and they are
independent there is no order defined for them (although this might change in
the future).

\subsection{Pre-run scans}

Most pre-run scans have the same priority, with a few exceptions, the most
important being \texttt{verifytext} to find out if a file is ASCII only, or if
there are any non-ASCII characters in the file. Since many of the scans
(including pre-run scans) only work on non-ASCII files it is important to find
out soon if a file contains only ASCII characters or not.

The order for pre-run scans is:

\begin{enumerate}
\item \texttt{checkXML}
\item \texttt{verifytext}
\item \texttt{verifyjava}
\item \texttt{verifyelf}, \texttt{verifysqlite3}
\item \texttt{verifyandroiddex}, \texttt{verifyandroidodex},
\texttt{verifyandroidresource}, \texttt{verifyandroidxml},
\texttt{verifycertificate}, \texttt{verifychromepak}, \texttt{verifyico},
\texttt{verifyihex}, \texttt{verifyjava}, \texttt{verifymessagecatalog},
\texttt{verifyresourcefork}, \texttt{verifyrsacertificate},
\texttt{verifyterminfo}, \texttt{verifytz}, \texttt{verifywebp},
\texttt{vimswap}
\end{enumerate}

\subsection{Unpack scans}

As a general rule of thumb: compressed formats are scanned last, while
simple containers that concatenate contents, or where the original content can
still be (partially) recognised, are scanned first.

An example of a container is TAR: content is simply concatenated without
compression. If the TAR archive would contain a file of a certain type (such as
a gzip compressed file) and the unpacker for that type is run first it will try
to carve it from the TAR file, blacklist the byte range, and the TAR unpacker
would not successfully run.

For the compressed files on the other hand the original content isn't visible
without unpacking so no other scans will pick it up and they can have a low
priority.

The order that is defined starts with \texttt{byteSwap}, a special unpacker
that is needed to unpack firmwares of certain devices, where a different kind
of flash chip is used, needing bytes in a firmware to be swapped first before
any other scan can be run.

Then the unpack scans for various container formats and file systems are run.
The order in which they appear is not fool proof: container files could be
embedded in container files with a lower priority, but BAT comes with
(hopefully) sane defaults to prevent this.

As a second to last step the unpack scans for compressed files where all data is
packed in such a way that the original content can't be seen without unpacking
are run.

Finally there are some scans that unpack text files (\texttt{base64}) or media
files. The \texttt{lzma} unpack scan also has the lowest priority because of
possibly many false positives.

The order of the unpack scans as defined in BAT 36 is:

\begin{enumerate}
\item \texttt{byteswap}
\item \texttt{tar}, \texttt{android-sparse}
\item \texttt{pdf\_unpack}, \texttt{iso9660}, \texttt{plf}, \texttt{wim}
\item \texttt{cramfs}, \texttt{ext2fs}, \texttt{ubi}
\item \texttt{ar}, \texttt{cpio}, \texttt{java\_serialized}, \texttt{romfs},
\texttt{rpm}, \texttt{upx}, \texttt{yaffs}
\item \texttt{exe}, \texttt{jffs2}, \texttt{squashfs}, \texttt{xar}
\item \texttt{7z}, \texttt{arj}, \texttt{bzip2}, \texttt{cab}, \texttt{chm},
\texttt{compress}, \texttt{gzip}, \texttt{installshield}, \texttt{intelhex},
\texttt{lrzip}, \texttt{lzip}, \texttt{lzo}, \texttt{minix}, \texttt{msi},
\texttt{pack200}, \texttt{rar}, \texttt{rzip}, \texttt{xz}, \texttt{zip},
\item \texttt{android-backup}, \texttt{base64}, \texttt{bmp}, \texttt{gif},
\texttt{ico}, \texttt{jpeg}, \texttt{lzma}, \texttt{ogg}, \texttt{otf},
\texttt{png}, \texttt{swf}, \texttt{ttf}, \texttt{woff}
\end{enumerate}

\subsection{Leaf scans}

There is currently only one explicit ordering: \texttt{kernelchecks} is run
before \texttt{identifier} because \texttt{identifier} depends on the result
of \texttt{kernelchecks}. For the rest the order of the leaf scans does not
matter.

\subsection{Aggregate scans}

Aggregate scans have a clear order. Reports and (most) images are generated at
the very end when all information is known. Other scans are mostly independent
of eachother, but are usually run before \texttt{versionlicensecopyright} to
prevent having to read big report pickles from disk.

The order for aggregate scans is:

\begin{enumerate}
\item \texttt{fixduplicates}
\item \texttt{prunefiles} (disabled by default)
\item \texttt{findduplicates}
\item \texttt{findlibs}, \texttt{findsymbols}, \texttt{copyright}, \texttt{file2package}
\item \texttt{kernelversions}
\item \texttt{versionlicensecopyright}
\item \texttt{passwords} (disabled by default), \texttt{shellinvocations}
\item \texttt{generateimages}, \texttt{generatereports}, \texttt{generatejson}, \texttt{searchlogins}
\end{enumerate}

\end{document}
